### 2020-08-17 2.background/background.tex

We establish the background necessary for the study of sparse signal detection and support recovery problems in this chapter.
Sections \ref{sec:risks} and \ref{sec:statistical-procedures} provide a refresher on the definitions of statistical risks and some commonly used statistical procedures.
Section \ref{sec:asymptotics} describes the asymptotic regime under which we analyze these procedures, and reviews the related literature in high-dimensional statistics.
We discuss in Section \ref{sec:risks-relations} the connections among the risk metrics, and point out some common fallacies.
The remaining sections collect the technical preparations for this text.
Section \ref{suppsec:AGG} defines an important class of error distributions which will be analyzed in detail in later chapters. 
And finally, Section \ref{subsec:RS} introduces the concepts of concentration of maxima, which plays a crucial role in the analysis of high-dimensional support recovery problems.

% prepare the definitions and technical results used in our analyses.

\section{Statistical risks}
\label{sec:risks}
\input{2.background/2.statistical_risks.tex}

\section{Statistical procedures}
\label{sec:statistical-procedures}
\input{2.background/2.statistical_procedures.tex}

\section{Related literature and our contributions}
\label{sec:asymptotics}
\input{2.background/2.asymptotics.tex}

\section{Relationships between the asymptotic risks}
\label{sec:risks-relations}
\input{2.background/2.risks_relations.tex}

\section{The asymptotic generalized Gaussian (AGG) models}
\label{suppsec:AGG}
\input{2.background/2.AGG.tex}

\section{Rapid variation and relative stability}
\label{subsec:RS}
\input{2.background/2.RV_and_RS.tex}