### 2020-08-17 3.phase_transitions/phase_transitions.tex

We study the fundamental limits of the signal detection and support recovery problems in the location models with independent Gaussian errors in this chapter.
Specifically, we derive the conditions under which the detection and support recovery problems succeed and fail in the sense of \eqref{eq:support-recovery-success} and \eqref{eq:support-recovery-failure}, in the additive error model 
\begin{equation} \label{eq:model-additive-Chapter3}
    x(i) = \mu(i) + \epsilon(i), \quad i=1,\ldots,p,
\end{equation}
where the errors $\epsilon(i)$'s are \ac{iid} standard Gaussians random variables.
We restrict our analysis to models with independent and identically distributed Gaussian errors for the moment.
Both the distributional assumption and the independence assumption will be relaxed in the following chapters.

As laid out in Section \ref{sec:asymptotics}, we work under the asymptotic regime where the problem dimension $p$ diverges to infinity.
Let the signal vectors $\mu = \mu_p$ have 
\begin{equation} \label{eq:signal-sparsity-additive}
    |S_p| = \left\lfloor p^{1-\beta} \right\rfloor, \quad \beta\in(0,1]
\end{equation}
non-zero entries, where $\beta$ parametrizes the problem sparsity.
The closer $\beta$ is to 1, the sparser the support $S$; conversely, when $\beta$ is close to 0, the support is dense with many non-null signals.
We consider one-sided alternatives \eqref{eq:global-test-one-sided}, and parametrize the range of the non-zero (and perhaps unequal) signals with
\begin{equation} \label{eq:signal-size-additive}
    \underline{\Delta} = \sqrt{2\underline{r}\log{p}}
    \le \mu(i) \le
    \overline{\Delta} = \sqrt{2\overline{r}\log{p}}, \quad \text{for all}\;\;i\in S_p,
\end{equation}
for some constants $0<\underline{r}\le\overline{r}\le+\infty$.

% \label{rmk:global-test-boundary}
The parametrization of signal sparsity \eqref{eq:signal-sparsity-additive} and signal sizes \eqref{eq:signal-size-additive} in the Gaussian model was first introduced in \citet{ingster1998minimax}, and later adopted by \cite{hall2010innovated}, \cite{cai2011optimal}, \cite{zhong2013tests}, \cite{cai2014optimal}, \cite{arias2017distribution1}, and numerous others for studying the signal detection problem in Gaussian location-scale models.
Similar scalings of sparsity and signal size are also used in, e.g., \cite{ji2012ups}, \cite{jin2014optimality}, \cite{butucea2018variable} to study the phase transitions of the support recovery problems under Gaussianity assumptions.

\section{Sparse signal detection problems}
\label{sec:global-tests}
\input{3.phase_transitions/3.global-tests.tex}

\section{Sparse signal support recovery problems}
\label{sec:additive-error-model-boundaries}
\input{3.phase_transitions/3.Gaussian-support-recovery.tex}

\section{Proofs}
\label{sec:proofs}
\input{3.phase_transitions/3.proofs-phase-transitions.tex}