### 2020-08-17 5.URS/URS.tex

In this chapter, we establish a complete characterization of URS for Gaussian arrays in terms of a simple condition on the covariance structures.
The condition is as follows.

% The proof of the main result is deferred until Section \ref{sec:proofs}.

% We define the class of dependence structures, referred to as uniformly decreasing dependence (UDD), for Gaussian arrays via their covariance matrices.

\begin{definition}[Uniformly decreasing dependence (UDD)] \label{def:UDD}
Consider a triangular array of jointly Gaussian distributed errors 
${\cal E} = \left\{\left(\epsilon_p(i)\right)_{i=1}^p, p = 1,2,\ldots\right\}$ 
with unit variances,
$$
\epsilon_p \sim \text{N}(0, \Sigma_p), \quad p=1,2,\ldots.
$$
The array ${\cal E}$ is said to be uniform decreasingly dependent (UDD) if 
for every $\delta>0$ there exists a finite $N(\delta)<\infty$, such that for every $i\in\{1,\ldots,p\}$, and $p\in\N$, we have
\begin{equation} \label{eq:UDD-definition}
    \Big|\left\{k\in\{1,\ldots,p\}:\Sigma_p(i,k)>\delta\right\}\Big| \le N(\delta)\quad \text{for all  } \delta>0.
\end{equation}
\end{definition}
That is, for every coordinate $i$, the number of elements which are more than $\delta$-correlated with $\epsilon_p(i)$ does not exceed $N(\delta)$.

Note that the bound in \eqref{eq:UDD-definition} holds uniformly in $i$ and $p$, and only depends on $\delta$.
Also observe that on the left-hand side of \eqref{eq:UDD-definition}, we merely count in each row of $\Sigma_p$ the number of exceedances of covariances (not their absolute values!) over level $\delta$.

\begin{remark} \label{rmk:choice-of-N(delta)}
Without loss of generality, we may require that $N(\delta)$ be a monotone non-increasing function of $\delta$, for we can take
$$
N(\delta) = \sup_{p,i} \Big|\{k:\Sigma_p(i,k)>\delta\}\Big|,
$$
which is non-increasing in $\delta$.
Definition \ref{def:UDD} therefore states that the array is UDD when $N(\delta)<\infty$ for all $\delta>0$.
\end{remark}

Observe that the UDD condition does not depend on the order of the coordinates in the error 
vector $\epsilon_p = (\epsilon_p(i))_{i=1}^p$.  Often times, however, the errors are thought of 
coming from a stochastic process indexed by time or space.  To illustrate the generality of the 
UDD condition, we formulate next a simple sufficient condition (UDD$^\prime$) that is easier to 
check in a time-series context.

\begin{definition}[UDD\,$^\prime$]\label{d:UDD-prime}
For $\epsilon_p \sim \mathrm{N}(0,\Sigma_p)$ with unit variances, an array ${\cal E} = \left(\epsilon_p(i)\right)_{i=1}^p$ is said to satisfy the UDD\,$^\prime$ condition if there 
exist:
\begin{enumerate}
    \item[(i)] permutations $l = l_p$ of $\{1,\ldots,p\}$, for all $p\in\N$, and
    \item[(ii)] a non-negative sequence $(r_n)_{n=1}^\infty$ converging to zero $r_n\to 0$, as $n\to\infty$,
\end{enumerate}
such that 
\begin{equation} \label{eq:weak-correlation}
    \sup_{p\in\N} |\Sigma_p\left(i',j'\right)| \le r_{|i-j|}.
\end{equation}
where $i' = l(i)$, $j' = l(j)$, for all $i,j\in\{1,\ldots,p\}$.
\end{definition}

\begin{remark}
Without loss of generality, we may also require that $r_n$ be non-increasing in $n$, for we can replace $r_n$ with the non-increasing sequence $r'_n = \sup_{m\ge n} r_m$.
\end{remark}

\begin{proposition} \label{prop:UDD-equivalent}
UDD\,$^\prime$ implies UDD.
\end{proposition}

\begin{proof}%[Proof of Proposition \ref{prop:UDD-equivalent}]
%The functions $N(\delta)$ and $r(n)=r_n$ are inverses of each other.
% {\bf $\text{UDD} \implies \text{UDD'}$:}
% If $N(0)<\infty$, then we can take 
% $$
% r_n = \underbrace{1, \ldots, 1}_{\lfloor N(0)/2\rfloor + 1}, 0, 0, \ldots.
% $$
% and recursively construct the permutations as follows.
% Start with any element

% We proof the contrapositive. 
% Suppose for all permutations and any sequence $r_n\to0$, there exists $i,j$ such that $\Sigma(i',j')>r_{|i'-j'|}$,
% then for any $\delta>0$ and any finite $M<\infty$, we can take $r_n$ to be a sequence of $M$ 1's followed by a $\delta$, i.e., 
% $$
% r_n = \underbrace{1, \ldots, 1}_{M+1}, \delta, \ldots.
% $$
% \fbox{Not true:}
% However, since there exists $i',j'$ such that $\Sigma(i',j')>r_{|i'-j'|}$, the set 
% $$
% S = l^{-1}\left(\left\{j',i'-N,\ldots,i'-1,i',i'+1,\ldots,i'+N\right\}\right)
% $$

% {\bf $\text{UDD'} \implies \text{UDD}$:} 
Since $r_n\to 0$, for any $\delta > 0$, there exists an integer 
$M = M(\delta)<\infty$ such that $r_n\le\delta$, for all $n\ge M$. 
Thus, by \eqref{eq:weak-correlation}, for every fixed 
$j' \in\{1,\ldots,p\}$, we can have $|\cov(\epsilon_p(k'),\epsilon_p(j'))| > \delta$,
only if $k'$ belongs to the set:
$$ 
 \left\{ k' \in \{1,\dots,p\} \, :\, j-M \le  k := l_p^{-1}(k') \le j+M \right\},
$$
where $j:= l_p^{-1}(j')$. That is, there are at most $2M+1<\infty$ indices  $k'\in\{1,\dots,p\}$, whose covariances with $\epsilon(j')$ may exceed $\delta$. 
Since this holds uniformly in $j'\in\{1,\ldots,p\}$, Condition UDD follows with 
$N(\delta) = 2M+1$.
\end{proof}

% \begin{definition}[Uniformly decreasing dependence (UDD)] \label{def:weak-dependence}
% Consider a triangular array of jointly Gaussian distributed errors $\left(\epsilon_p(j)\right)_{j=1}^p$ with unit variances, $\epsilon_p \sim \mathcal N(0,\Sigma_p)$. The array ${\cal E}$ is said to be uniform decreasingly dependent (UDD) with rate $(r_n)_{n=1}^\infty$ if 
% \begin{equation} \label{eq:weak-correlation}
%     \sup_p |\Sigma_p(i,j)| \le r_{|i-j|}
% \end{equation}
% such that $r_n\to 0$, as $n\to\infty$.
% \end{definition}
% 
% \begin{remark} \label{rmk:UDD-equivalent}
% In situations where there is no natural ordering of the components, it is sufficient that a permutation of the vector $\epsilon$ in its coordinates satisfy the requirements above. 
% In fact, the UDD condition can be equivalently stated as follows: for any $\delta>0$, and any coordinate $j\in\{1,\ldots,p\}$, there are at most $N(\delta)<\infty$ coordinates whose covariances with $\epsilon(j)$ exceed $\delta$; here $N(\delta)$ is a deterministic function independent of $p$.
% $N(\delta) \to 1$ as $\delta \to 0$.
% \end{remark}

We now state the main result of this section: a Gaussian sequence is URS if and only if it is UDD.
The URS condition essentially requires that the dependencies decay in a uniform fashion, the rate at which dependence decay does \emph{not} matter.

\begin{theorem} \label{thm:Gaussian-weak-dependence}
Let ${\cal E}$ be a Gaussian triangular array with standard normal marginals.  
The array ${\cal E}$ has uniformly relatively stable (URS) maxima if and only if it is uniformly decreasing dependent (UDD).
\end{theorem}

% The proof of Theorem \ref{thm:Gaussian-weak-dependence} is given in Section \ref{sec:proofs}.

Specifically, for stationary Gaussian arrays, we have the following corollary.

\begin{corollary} \label{cor:stationary-Gaussian-errors}
Let ${\cal E} = \{\epsilon_p(i) = Z(i)\}$ for a stationary Gaussian time series ${\cal Z} = \{Z(i)\}$.
Then $\cal E$ is {URS} if and only if the autocovariance function $\cov(Z(k), Z(0))\to 0$, as $k\to\infty$.
\end{corollary}

Corollary \ref{cor:stationary-Gaussian-errors} follows by Theorem \ref{thm:Gaussian-weak-dependence} and the observation that UDD is equivalent to vanishing autocovariance of $\cal Z$.
A slightly weaker form of the ``if'' part was established in Theorem 3 of \cite{berman1964limit}.

Returning again to the study of support recovery problems, Theorem \ref{thm:Gaussian-weak-dependence} and the necessary condition for exact support recovery in Theorem \ref{thm:necessary} yields the following result.

\begin{corollary} \label{cor:weakly-dependent-errors}
For UDD Gaussian errors, the result in Theorem \ref{thm:necessary} holds.
\end{corollary}

As a counterpart to Remark \ref{rmk:dependence-assumptions}, we demonstrate the tightness of the dependence conditions in Theorem \ref{thm:necessary}.
Specifically, we demonstrate that if the URS dependence condition is violated, then it may be possible to recover the support of weaker signals below the boundary.

\begin{example} \label{exmp:counter-example}
Suppose ${\cal E} = \left(\epsilon_p(i)\right)_{i=1}^p$ is Gaussian, and is comprised of $\lfloor p^{1-\beta}\rfloor$ blocks, each of size at least $\lfloor p^\beta \rfloor$; 
let the elements of each block have correlation 1, and let elements from different blocks be independent. 
If $\underline{r} \ge 4(1-\beta)$, then the procedure $\widehat{S} = \big\{i:x(i)>\sqrt{2(1-\beta)\log{p}}\big\}$ yields $\mathbb P[\widehat{S} = S] \to 1$. 
This requirement on signal size is strictly weaker than that of the strong classification boundary, since $4(1-\beta) < (1 + \sqrt{1-\beta})^2$ on $\beta\in(0,1)$.
\end{example}

The above example shows that if the correlations of the Gaussian errors do not decay in a uniform fashion (UDD fails), then we can do substantially better in terms of support recovery.
The claims in the example are verified in Section \ref{suppsec:proofs}, while numerical simulations of this example can be found in Section \ref{suppsec:numerical}.

\medskip

We conclude with a brief discussion on the relationships between UDD and other dependence conditions in the context of extreme value theory, before proceeding to the proof of Theorem \ref{thm:Gaussian-weak-dependence}.

Suppose that the array of errors  ${\cal E}$ comes from a stationary Gaussian time series $\epsilon(i),\ i\in \mathbb{N}$, with auto-covariance $r_p=\cov(\epsilon(i+p),\epsilon(i))$. 
One is interested in the asymptotic behavior of the maxima $M_p:=\max_{i=1,\dots,p} \epsilon(i)$.

In this setting, the Berman's condition, introduced in \cite{berman1964limit}, requires that
\begin{equation} \label{eq:Berman}
    r_p \log p \to 0,\ \ \mbox{ as }p\to\infty.
\end{equation}
This condition entails that 
\begin{equation}
    \label{eq:Gauss-max-in-distribution}
  a_p (M_p - b_p) \stackrel{d}{\longrightarrow } Z,\  \ \mbox{ as }p\to\infty,
\end{equation}
with the Gumbel limit distribution $\mathbb P [Z\le x] = \exp\{-e^{-x}\},\ x\in \mathbb R$, 
where 
$$
a_p = \sqrt{2\log p},\quad b_p  = \sqrt{2\log p} - \frac{1}{2}\left(\sqrt{2\log p}\right)^{-1}\left(\log \log (p) + \log(4\pi)\right),
$$ 
are {\em the same} centering and normalization sequences
as in the case of iid $\epsilon(i)$'s.  
Berman's condition is one of the weakest dependence conditions  in the literature for which this result holds. See, e.g., Theorem 4.4.8 in \cite{embrechts2013modelling}, where \eqref{eq:Berman} is described as ``very weak''.

For dependence conditions weaker than \eqref{eq:Berman}, the sequences of normalizing and centering constants in \eqref{eq:Gauss-max-in-distribution} are {\em different} from the iid case, and the corresponding limit is no longer Gumbel; see, for example, Theorems 6.5.1 and  6.6.4 in \cite{leadbetter2012extremes}, and \cite{mccormick1976weak}. 
% In particular, \cite{mccormick1976weak} derived the normalizing constants when both $r_p\to 0$ monotonically and  $r_p \log p \to \infty$ monotonically, as $p\to\infty$. 
% In this case, convergence in distribution still takes place, with the maxima concentrating along a sequence asymptotic to \eqref{eq:quantiles}.

On the other hand, in our high dimensional support estimation context, the notion of relative stability is sufficient and more natural than the finer notions of distributional convergence.
If one is merely interested in the asymptotic relative stability of the Gaussian maxima, then Berman's condition can be relaxed significantly (see also, Theorem 4.1 of \cite{berman1964limit}). 
Observe that by Proposition \ref{prop:UDD-equivalent},  the Berman condition \eqref{eq:Berman} implies UDD and hence relative stability (Theorem \ref{thm:Gaussian-weak-dependence}), i.e., 
\begin{equation} \label{eq:Gaussian-URS}
  \frac{1}{b_p} M_p \stackrel{\mathbb P}{\to} 1,\quad\mbox{as}\quad p\to\infty.
\end{equation}
This {\em concentration of maxima} property can be readily deduced from \eqref{eq:Gauss-max-in-distribution}, since $a_p b_p \sim 2\log(p) \to \infty$ as $p\to\infty$.
Theorem \ref{thm:Gaussian-weak-dependence} shows that \eqref{eq:Gaussian-URS} holds if the much weaker uniform dependence condition UDD holds. 
Note that our condition is coordinate free --- neither monotonicity of the sequence $r_p$ nor stationarity of the underlying array is required.

% The method of proof is also very different from the results on distributional convergence in the references mentioned above.

\medskip

The rest of this chapter is devoted to the proof of the main result, i.e., Theorem \ref{thm:Gaussian-weak-dependence}. 
We first introduce a key lemma regarding the structure of correlation matrix of high-dimensional random variables.
The proof uses a surprising, yet elegant application of Ramsey's Theorem from the study of combinatorics.
%; this application, and its consequences in high-dimensional probability, are presented in Section \ref{subsec:Ramsey}.
The `only if' part of Theorem \ref{thm:Gaussian-weak-dependence} follows from this lemma, in Section \ref{sec:URS=>UDD}.

The proof of the `if' part is detailed in Section \ref{sec:UDD=>URS}.
The arguments there was recently extended to establish bounds on the rate of concentration of maxima in \cite{kartsioukas2019rate}; see also, \cite{tanguy2015some} and references therein for related work on this topic.

\section{Ramsey's Coloring theorem and structure of correlation matrices} 
\label{sec:Ramsey}
\input{5.URS/5.proofs-Ramsey.tex}

\section{URS implies UDD (`only if' part of Theorem \ref{thm:Gaussian-weak-dependence})} 
\label{sec:URS=>UDD}
\input{5.URS/5.proofs-URS-UDD.tex}

\section{UDD implies URS (`if' part of Theorem \ref{thm:Gaussian-weak-dependence})} 
\label{sec:UDD=>URS}
\input{5.URS/5.proofs-UDD-URS.tex}

\section{Numerical illustrations of exact support recovery under dependence}
\label{sec:URS-numerical}
\input{5.URS/5.numerical.tex}