### 2020-08-17 6.GWAS/6.odds-and-power.tex

We return to the application of association screenings for categorical variables, and put the results in the previous section to use.
In particular, we focus on the exact-approximate support recovery problem, and demonstrate the consequences of its phase transition (Theorem \ref{thm:chi-squared-exact-approx-boundary}) in genetic association studies.

In order to do so, we must first connect the concept of ``statistical signal size'' $\lambda$ with some key quantities in association tests.
While ``signal size'' likely sounds foreign to most practitioners, it is intimately linked with the concept of ``effect sizes'' --- or odds ratios --- in association studies, which are frequently estimated and reported in GWAS catalogs.
We characterize the relationship between the two quantities in the special, but fairly common case of association tests on 2-by-2 contingency tables in Section \ref{sec:odds-and-power}.

% Unlike in additive models where the parameter $\mu$ has the interpretation of signal-to-noise ratios, the meaning of the signal sizes $\lambda$ in chi-square model is perhaps not as transparent.

Consider a 2-by-2 multinomial distribution with marginal probabilities of phenotypes $(\phi_1, \phi_2)$ and genotypes $(\theta_1, \theta_2)$.
The \emph{probability} table (as opposed to the table of multinomial \emph{counts} in the introduction) is as follows.
\begin{center}
    \begin{tabular}{cccc}
    \hline
    & \multicolumn{2}{c}{Genotype} \\
    \cline{2-3}
    Probabilities & Variant 1 & Variant 2 & Total by phenotype \\
    \hline
    Cases & $\mu_{11}$ & $\mu_{12}$ & $\phi_1$ \\
    Controls & $\mu_{21}$ & $\mu_{22}$ & $\phi_2$ \\
    Total by genotype & $\theta_1$ & $\theta_2$ & 1 \\
    \hline
    \end{tabular}
\end{center}
The odds ratio (i.e., ``effect size'') is defined as the ratio of the phenotype frequencies between the two genotype variants,
\begin{equation} \label{eq:odds-ratio}
    \text{R} := \frac{\mu_{11}}{\mu_{21}}\Big/\frac{\mu_{12}}{\mu_{22}}
    = \frac{\mu_{11}\mu_{22}}{\mu_{12}\mu_{21}}.
\end{equation}
The multinomial distribution is fully parametrized by the trio $(\theta_1, \phi_1, R)$.
Odds ratios further away from 1 indicate greater contrasts between the probability of outcomes.
Independence between the genotypes and phenotypes would imply an odds ratio of one, and hence $\mu_{jk} = \phi_j\theta_k$, for all $j,k \in\{1,2\}$.

% When data are sampled from the multinomial distribution, the chi-square test defined in \eqref{eq:chisq-statistic} is asymptotically equivalent to tests including, e.g., the likelihood ratio test and Welch's t-test, both in terms of level and power \cite{ferguson2017course,gao2019upass}.
For a sequence of local alternatives $\mu^{(1)}, \mu^{(2)}, \ldots$, such that $\sqrt{n}(\mu^{(n)}_{jk} - \phi_j\theta_k)$ converges to a constant table $\delta = (\delta_{jk})$, the chi-square test statistics converge in distribution to the non-central chi-squared distribution with non-centrality parameter 
\begin{equation*}
    \lambda = \sum_{j=1}^2 \sum_{k=1}^2 {\delta_{jk}^2}/{(\phi_j\theta_k)}.
\end{equation*}
See, e.g., \cite{ferguson2017course}.
Hence, for large samples from a fixed distribution $(\mu_{ij})$, the statistic is well approximated by a $\chi^2_1(\lambda)$ distribution, where
\begin{equation} 
\lambda = n\sum_{j=1}^2 \sum_{k=1}^2 \frac{(\mu_{jk} - \phi_j\theta_k)^2}{\phi_j\theta_k}.
\end{equation}
%Since $\lambda$ is linear in the number of samples $n$, 
% Power of association tests at $\alpha$ level is approximately $\P[\chi^2_{\nu}(\lambda)>\chi^2_{\nu,\alpha}]$, where $\chi^2_{\nu,\alpha}$ is the upper $\alpha$-quantile of a central Chi-squared distribution.
Power calculations therefore only depend on the $\mu_{jk}$'s through $\lambda=nw^2$, where we define 
\begin{equation} \label{eq:signal-size-chisq}
    w^2:=\lambda/n
\end{equation} 
to be the \emph{signal size per sample}. 
Statistical power would be increasing in $w^2$ for fixed sample sizes.

The next proposition states that the statistical signal size per sample can be parametrized by the odds ratio and the marginals in the probability table.

\begin{proposition} \label{prop:signal-size-odds-ratio}
Consider a 2-by-2 multinomial distribution with marginal distributions $(\phi_1, \phi_2 = 1-\phi_2)$ and $(\theta_1, \theta_2=1-\theta_1)$.
Let signal size $w^2$ be defined as in \eqref{eq:signal-size-chisq}, and odds ratio $\text{R}$ be defined as in \eqref{eq:odds-ratio}. 
If $R=1$, we have $w^2 = 0$; if $R\in(0,1)\cup(1,+\infty)$, then we have
\begin{equation} \label{eq:signal-size-odds-ratio}
    w^2(\text{R}) =
    \frac{1}{4A(\text{R}-1)^2}\left(B+CR-\sqrt{(B+CR)^2-4A(R-1)^2}\right)^2,
\end{equation}
where $A = \phi_1\theta_1\phi_2\theta_2$, $B = \phi_1\theta_1+\phi_2\theta_2$, and $C = \phi_1\theta_2+\phi_2\theta_1$.
\end{proposition}

Proposition \ref{prop:signal-size-odds-ratio} is derived in Section \ref{subsec:proof-signal-size-odds-ratio}.

To understand Proposition \ref{prop:signal-size-odds-ratio}, we illustrate Relation \eqref{eq:signal-size-odds-ratio} for selected values of marginals $\theta_1$ and $\phi_1$ in Figure \ref{fig:signal-vs-odds}.
Observe in the figure that an odds ratio further away from one corresponds to stronger statistical signal per sample, ceteris paribus.
However, this ``valley'' pattern is in general not symmetric around 1, except for balanced marginal distributions ($\phi_1=1/2$ or $\theta_1=1/2$).
While the odds ratio $R$ can be arbitrarily close to 0 or diverge to $+\infty$ for any marginal distribution, the signal sizes $w^2$ are bounded from above by constants that depend only on the marginals.
% This is quantified in the next corollary.

\begin{figure}
      \centering
      \includegraphics[width=0.49\textwidth]{pics/singal-vs-odds-p05}
      \includegraphics[width=0.49\textwidth]{pics/singal-vs-odds-p0333}            
      \caption{Signal sizes per sample $w^2$ as functions of odds ratios in 2-by-2 multinomial distributions for selected genotype marginals in balanced (left) and unbalanced (right) designs; see Relation \eqref{eq:signal-size-odds-ratio} in Proposition \ref{prop:signal-size-odds-ratio}. For given marginal distributions, extreme odds ratios imply stronger statistical signals at a given sample size. However, the signal sizes are bounded above by constants that depend on the marginal distributions; see Relations \eqref{eq:signal-size-upper-bound-1} and \eqref{eq:signal-size-upper-bound-2}. % Unbalanced marginal distributions -- or rare variants -- lead to smaller signal sizes at a given odds ratio.
      } 
      \label{fig:signal-vs-odds}
\end{figure}

\begin{corollary} \label{cor:signal-limits-OR}
The signal size as a function of the odds ratio $w^2(R)$ is decreasing on $(0,1)$ and increasing on $(1,\infty)$, with limits
\begin{equation} \label{eq:signal-size-upper-bound-1}
    \lim_{\text{R}\to0_+} w^2(\text{R}) = \min\left\{\frac{\phi_1\theta_1}{\phi_2\theta_2}, \frac{\phi_2\theta_2}{\phi_1\theta_1}\right\},
\end{equation}
and
\begin{equation} \label{eq:signal-size-upper-bound-2}
    \lim_{\text{R}\to+\infty} w^2(\text{R}) = \min\left\{\frac{\phi_1\theta_2}{\phi_2\theta_1}, \frac{\phi_2\theta_1}{\phi_1\theta_2}\right\}.
\end{equation}
\end{corollary}
% Proof of Corollary \ref{cor:signal-limits-OR} is found in Appendix \ref{sec:proof-signal-size-odds-ratio}.

Corollary \ref{cor:signal-limits-OR} immediately implies that balanced designs with roughly equal number of cases and controls are not necessarily the most informative.

For example, in a study where a third of the recruited subjects carry the genetic variant positively correlated with the trait (i.e., $\theta_1=1/3$), an unbalanced design with $\phi_1=1/3$ would maximize $w^2$ at large odds ratios.
This unbalanced design is much more efficient compared to, say, a balanced design with $\phi_1=1/2$.
In the first case, we have $w^2\to1$ as $R\to\infty$; whereas in the second design, $w^2<1/2$ no matter how large $R$ is.
This difference can also be read by comparing the dashed curve ($\theta_1=1/3$, $\phi_1=1/2$) in the left panel of Figure \ref{fig:signal-vs-odds}, with the solid curve ($\theta_1=1/3$, $\phi_1=1/3$) in the right panel of Figure \ref{fig:signal-vs-odds}.