### 2020-08-17 SB_draft.tex

% Use the University of Michigan thesis class.
\documentclass[thesis]{./tex/thesis-umich}

% Use natbib for citation styling
\usepackage{natbib}
%\bibliographystyle{abbrvnat}
\bibliographystyle{apa}
\setcitestyle{authoryear,open={(},close={)}}

%
% My definitions:
%
\def\bbN{{\mathbb N}}
\def\bbZ{{\mathbb Z}}
\def\bbR{{\mathbb R}}
\def\bbC{{\mathbb C}}

\def\P{{\mathbb P}}
\def\E{{\mathbb E}}

\newcommand{\var}{{\rm {Var}}}
\newcommand{\cov}{{\rm {Cov}}}

\def\N{\mathbb N}
\def\R{\mathbb R}
\def\Z{\mathbb Z}
\usepackage{bm}
\renewcommand\vec[1]{{\bm #1}}

\def\alert#1{\fbox{#1}}
\def\what{\widehat}
\def\noi{\noindent}
\def\MSE{{\rm m.s.e.}}
% \newcommand\comment[1]{{}}
\def\refeq#1{(\ref{e:#1})}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
\newcommand\version[2]{{#1}} % The short version of the paper
%\newcommand\version[2]{{#2}} % The long version of the paper

%\startlocaldefs
\numberwithin{equation}{chapter}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
% my defs
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{assumption}{Assumption}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
%\endlocaldefs

% Title of the thesis
\title{Concentration of Maxima and Fundamental Limits in High-Dimensional  Testing and Inference}
% \title{Phase Transitions in High-dimensional Testing Problems and Concentration of Maxima}

% Author name
\author{Zheng Gao\\Stilian Stoev}

% Email
\email{gaozheng@umich.edu\\sstoev@umich.edu}

% ORCID
\ORCID{xxxx-xxxx-xxxx-xxxx}

% Department
\department{Statistics}

% Year of completion
\year=2020

% Frontispiece
\frontispiece{
\includegraphics[width=5.5in]{./pics/frontispiece.jpg}\\
\vspace{1in}
\begin{flushleft}
{%\fontfamily{pbk}\selectfont 
For what is a man, what has he got? %\\
%If not himself, then he has naught.
}
\end{flushleft}
}

% Default style for front pages
\frontpagestyle{1}

% Dedication
\dedication{ %
\centering
}

% Acknowledgments
\acknowledgments{ %
}

% This command sets the width of the acknowledgments area as a fraction
% of the total width of the text area.
\acknowledgmentswidth{0.8}

% Preface
% \preface[2]{ %
% This is a preface.}

% Commands to hide or show lists of figures, tables, etc.
\hidelistoftables
% \showlistofprograms
% \showlistofappendices

% Definition of any abbreviations used.
\abbreviations{
 \acro{AGG}{asymptotically generalized Gaussian}
 \acro{BH}{Benjamini-Hochberg}
 \acro{CDF}{cumulative distribution function}
 \acro{FDR}{false discovery rate}
 \acro{FNR}{false non-discovery rate}
 \acro{FWER}{family-wise error rate}
 \acro{FWNR}{family-wise non-discovery rate}
 \acro{GWAS}{genome-wide association studies}
 \acro{HC}{higher-criticism}
 \acro{iid}{independent and identically distributed}
 \acro{LR}{likelihood-ratio}
 \acro{RS}{relatively stable/relative stability}
 \acro{SNP}{single-nucleotide polymorphisms}
 \acro{URS}{uniform relatively stable/uniform relative stability}
}

% Some abstract text
\abstract{
This text presents a collection of new results and recent developments on the phase transition phenomena in sparse signal problems, and makes the connection to the forces behind these statistcal results --- a probablistic phenomena known as concentration of maxima.
We begin with idealized models where high-dimensional signals are observed with independent additive Gaussian noise.
Under suitable parametrization of the signal sparsity and signal sizes, 
we characterize several new phase transition phenomena in high-dimensional additive error models, and derive the signal sizes necessary and sufficient for statistical procedures to simultaneously control false discovery (in terms of Type I error, {FDR}, or {FWER}) and missed detection (in terms of Type II error, {FNR}, or {FWNR}) in large dimensions.
% Several new phase transition phenomena are characterized.
Specifically, we show that, depending on the signal sparsity and signal sizes, the sum of the statistical risks for false discovery  and missed detection either can be controlled at very low levels in high dimensions, or must be at least one in the limit, for all thresholding procedures.

The phase transition in the case of family-wise type I and type II error control is further generalized to cover non-Gaussian and dependent observations.
We show that the sharp zero-one law continues to hold under a very broad class of error dependence structures, and for a very broad class of error distributions with light, rapidly varying, tails.
The key to this important generalization is a certain {concentration of maxima} phenomenon known as relative stability. 
We demystify the dependence conditions of the phase transition by providing a complete characterization of the relative stability concept for Gaussian triangular arrays in terms of their correlation structures.
% The proof uses classic Sudakov-Fernique and Slepian lemma arguments along with a curious application of Ramsey's coloring theorem.

Finally, motivated by marginal screenings of categorical variables, we study high-dimensional multiple testing problems where test statistics have approximate chi-square distributions.
Phase transitions are established for support recovery problems in high-dimensional chi-square models.
Remarkably, the degree-of-freedom parameters in the chi-square distributions do not affect the boundaries in all phase transitions.
% Several well-known procedures are shown to attain these boundaries.
We also elucidate on the notion of signal sizes in association tests by characterizing its relationship with marginal frequencies, odds ratio, and sample sizes in $2\times2$ contingency tables. 
This allows us to illustrate an interesting manifestation of the phase transition phenomena in % \ac{GWAS},
genome-wide association studies (GWAS), 
and explain some long-standing empirical observations on the discoverability of signals in GWAS.
As an auxiliary result, we show that, perhaps surprisingly, given total sample sizes, balanced designs in such association studies rarely deliver optimal power. % for detecting rare genetic variants with high odds ratios.
% We provide explicit formulas for power calculations, as well as software tools for finding the optimal study designs.
% The phase transition phenomena is demonstrated with numerical simulations, and with empirical evidence from GWAS.
% data from the NHGRI-EBI GWAS catalog.
}
%\hideabstractpagenumber

%% DOCUMENT AREA
\begin{document}

\chapter{Introduction}
\label{chap:intro}
\input{./1.intro/introduction.tex}

\chapter{Techincal Preparations and Background}
\label{chap:background}
\input{./2.background/background.tex}

\chapter{The Phase Transition Phenomena in Independent Gaussian Error Models} 
\label{chap:phase-transitions}
\input{./3.phase_transitions/phase_transitions.tex}

\chapter{Exact Support Recovery and Minimax Optimality} 
\label{chap:exact-support-recovery}
\input{./4.exact_support_recovery/exact_support_recovery.tex}

\chapter{Characterization of Uniform Relative Stability for Gaussian Arrays}
\label{chap:URS}
\input{./5.URS/URS.tex}

\chapter{The Phase Transition Phenomena in Genome-wide Association Studies} 
\label{chap:GWAS}
\input{./6.GWAS/GWAS.tex}

% \appendix
% \chapter{U-PASS: A Software for Unified Power Analysis and Forensics for Qualitative Traits in Genetic Association Studies} \label{app:UPASS}
% \addcontentsline{toc}{chapter}{Appendix~\ref{app:UPASS}: U-PASS: A Software for Unified Power Analysis and Forensics for Qualitative Traits in Genetic Association Studies}
% \input{./appendix/UPASS.tex}

% \chapter{Using Appendices} \label{app:appendix2}

% Give this command the relative path to the .bib file.
\bibliography{./tex/thesis}

\end{document}