### 2020-08-17 appendix/A.power_analysis.tex

We briefly review the main steps of a typical power analysis for genetic association studies.

\medskip
{\bf 1. Disease model specifications.}
A typical power analysis for genetic association studies begin by specifying an alternative hypothesis through a disease model (dominant, recessive, multiplicative, additive, etc.), which assumes:

\begin{itemize}
    \item The genotype relative risks (GRR).
    \item Risk allele frequency in the general population ($p$).
    \item Disease prevalence in the general population (Prev).
\end{itemize}

The disease model and parameters determine the joint distribution of the genotypes and phenotyes {\it in the population}, shown in the following table.

\begin{center}
    \begin{tabular}{cccc}
    \hline
    & \multicolumn{3}{c}{Risk allele copies} \\
    \cline{2-4}
    Population Prob. & 0 copies & 1 copy & 2 copies \\
    \hline
    Cases & $\pi_{10}$ & $\pi_{11}$ & $\pi_{12}$ \\
    Controls & $\pi_{20}$ & $\pi_{21}$ & $\pi_{22}$ \\
    \hline
    \end{tabular}
\end{center}

In the disease models, the conditional probabilities of having the disease, given the risk allele copy numbers, satisfy the following relations,

\begin{equation} \label{eq:GRR}
    \frac{\pi_{10}}{\pi_{10} + \pi_{20}} : \frac{\pi_{11}}{\pi_{11} + \pi_{21}} : \frac{\pi_{12}}{\pi_{12} + \pi_{22}}
    = \begin{cases}
    1 : \text{GRR} : \text{GRR}^2, &\text{Multiplicative} \\
    1 : \text{GRR} : 2\times\text{GRR}-1, & \text{Additive} \\
    1 : \text{GRR} : \text{GRR}, & \text{Dominant} \\
    1 : 1 : \text{GRR}, & \text{Recessive}
\end{cases}
\end{equation}
where GRR is strictly greater than 1 under the alternative, and equal to 1 under the null hypothesis.

The disease prevalence determines the sum of the probabilities of cases in the population,
\begin{equation} \label{eq:prev}
    \pi_{10} + \pi_{11} + \pi_{12} = \text{Prev}.
\end{equation}

The risk allele frequency in the general population, $p$, assuming Hardy-Weinberg equilibrium, satisfies 
\begin{equation} \label{eq:RAF-pop}
    \pi_{10} + \pi_{20} = (1-p)^2, \quad \pi_{11} + \pi_{21} = 2p(1-p), \quad \pi_{12} + \pi_{22} = p^2.
\end{equation}

The population probabilities are determined by the disease model and its parameters (GRR, Prev, and $p$).
The six unknowns  $(\pi_{10},\ldots,\pi_{22})$ and are solved for using the six equations above: two from Relation \eqref{eq:GRR}, one from \eqref{eq:prev}, and three from \eqref{eq:RAF-pop}.

\bigskip
{\bf 2. Sampling adjustments.}
Next, the probabilities of observing each genotype-phenotype combination are adjusted according the number of cases and controls recruited \emph{in the studies}, where the sample sizes are specified with the number of cases ($n_1$) and controls ($n_2$), or equivalently, the fraction of cases ($\phi$) and total number of subjects ($n$).

\begin{center}
    \begin{tabular}{cccc}
    \hline
    & \multicolumn{3}{c}{Risk allele copies} \\
    \cline{2-4}
    Prob. in study & 0 copies & 1 copy & 2 copies \\
    \hline
    Cases & $\pi_{10}\frac{\phi}{\text{Prev}}$ & $\pi_{11}\frac{\phi}{\text{Prev}}$ & $\pi_{12}\frac{\phi}{\text{Prev}}$ \\
    Controls & $\pi_{20}\frac{1-\phi}{1-\text{Prev}}$ & $\pi_{21}\frac{1-\phi}{1-\text{Prev}}$ & $\pi_{22}\frac{1-\phi}{1-\text{Prev}}$ \\
    \hline
    \end{tabular}
\end{center}

As an example, if $\phi > \text{Prev}$, the probabilities are adjusted to account for over-sampling of cases.
Conversely, adjustments are needed to account for under-sampling of cases.

The relative frequencies of allele type-phenotype combinations {\it in the study} are then calculated as follows.

\begin{center}
    \begin{tabular}{ccc}
    \hline
    & \multicolumn{2}{c}{Allele variant} \\
    \cline{2-3}
    Prob. in study & Risk allele & Non-risk allele \\
    \hline
    Cases & $\phi\left(\frac{\pi_{12}}{\text{Prev}}+\frac{\pi_{11}}{2\times\text{Prev}}\right)$ & $\phi\left(\frac{\pi_{11}}{2\times\text{Prev}}+\frac{\pi_{10}}{\text{Prev}}\right)$ \\
    Controls & $(1-\phi)\left(\frac{\pi_{22}}{1-\text{Prev}}+\frac{\pi_{21}}{2(1-\text{Prev})}\right)$ & $(1-\phi)\left(\frac{\pi_{21}}{2(1-\text{Prev})}+\frac{\pi_{20}}{1-\text{Prev}}\right)$ \\
    \hline
    \end{tabular}
\end{center}

This final table corresponds to the probabilities underlying the $2\times2$ multinomial counts that we introduced in Chapter \ref{chap:intro}.
We denote the relative frequencies of allele type-phenotype combinations with $\mu = (\mu_{11}, \mu_{12}, \mu_{21}, \mu_{22})$, as we did in Chapter \ref{chap:GWAS}.

\begin{center}
    \begin{tabular}{cccc}
    \hline
    & \multicolumn{2}{c}{Allele variant} & \\
    \cline{2-3}
    Prob. in study & Risk allele & Non-risk allele & Total by phenotype \\
    \hline
    Cases & $\mu_{11}$ & $\mu_{12}$ & $\phi = \mu_{11} + \mu_{12}$ \\
    Controls & $\mu_{21}$ & $\mu_{22}$ & $1-\phi = \mu_{21} + \mu_{22}$ \\
    \hline
    \end{tabular}
\end{center}

\bigskip
{\bf 3. Power calculations.}
Finally, the power of an statistical test is calculated as the probability of (a correct) rejection, assuming that the data (i.e., tabulated counts of the allele type-phenotype combinations) follow a multinomial or binomial distribution with sample size $2n$, since each individual has a pair of alleles.

These steps form the basis of the calculations implemented in the most existing tools, including the GAS calculator \citep{johnson2017gas}.

Some common association tests include the likelihood ratio test, Pearson's chi-square test, tests of zero slope coefficient in logistic regressions, as well as t-tests for equal proportions.
Although not explicitly stated, the GAS calculator assumes the test of association to be Welch's t-test.
In principal, power analysis has to be tailored to the association test used.
Fortunately, many of these tests are asymptotically equivalent in terms of power \citep{ferguson2017course, gao2019upass}, and results of the power approximation applies to all asymptotically equivalent tests.

% \footnote{Caveat: The final step in the power calculations does not apply to association tests performed directly on the 2-by-3 contingency tables of phenotype-genotype combinations (e.g., the Cochran-Armitage test).Therefore, the GAS calculator should not be used if these tests are to be applied.}.