### 2020-08-17 appendix/UPASS.tex

We introduce the software we developed for the power analysis of genetic association studies of qualitative traits in this appendix.
Section \ref{sec:power-analysis-overview} reviews the typical process of study planning employed by geneticists.
Disease models are often used to specify the distribution of observations under the alternative hypothesis in power analysis. 
In Section \ref{sec:disease-models}, we revisit the use of these disease models and argue for an alternative way of model parametrization that is better suited to conducting systematic reviews and confirmatory studies.
Section \ref{sec:user-guide} provides three example usages of the software.
Instructions for downloading, installing, launching, and terminating the software can be found in Section \ref{sec:UPASS-misc}.

\section{Power analysis in genetic association studies}
\label{sec:power-analysis-overview}
\input{./appendix/A.power_analysis.tex}

\section{Specification of alternatives in power analysis}
\label{sec:disease-models}
\input{./appendix/A.alternative_specification.tex}

\section{Use case illustrations}
\label{sec:user-guide}
\input{./appendix/A.use_case_illustrations.tex}

\section{Download, installation, and usage}
\label{sec:UPASS-misc}
\input{./appendix/A.setup.tex}