### 2020-08-18 acronyms.tex

\extrachap{Acronyms}

 \begin{description}[CABR]
  \item[AGG]{asymptotically generalized Gaussian}
  \item[BH]{Benjamini-Hochberg}
 \item[CDF]{cumulative distribution function}
 \item[FDR]{false discovery rate}
 \item[FNR]{false non-discovery rate}
 \item[FWER]{family-wise error rate}
 \item[FWNR]{family-wise non-discovery rate}
 \item[GWAS]{genome-wide association studies}
 \item[HC]{higher-criticism}
 \item[iid]{independent and identically distributed}
 \item[LR]{likelihood-ratio}
 \item[RS]{relatively stable/relative stability}
 \item[SNP]{single-nucleotide polymorphisms}
 \item[URS]{uniform relatively stable/uniform relative stability}
 \end{description}