### 2020-08-21 6.GWAS/GWAS.tex

We investigate the fundamental limits of multiple testing problems in high-dimensional chi-square models, and in genome-wide association studies as introduced in Section \ref{sec:motivation-chisq}.

In Section \ref{sec:chisq-boundaries}, we shall establish the phase transitions of the sparse chi-square model \eqref{eq:model-chisq}.
Recall that in large-scale screening studies where a large number of association tests are conducted, resulting statistics may be approximated by
\begin{equation} \label{eq:model-chisq-Chapter6}
    %x(i) \distras{\mathrm{ind.}} \chi_\nu^2\left(\lambda(i)\right), \quad i=1,\ldots,p.
    x(i) \sim \chi_\nu^2\left(\lambda(i)\right), \quad i=1,\ldots,p,
\end{equation}
where $\chi_\nu^2\left(\lambda(i)\right)$ is a chi-square distributed random variable with $\nu$ degrees of freedom and non-centrality parameter $\lambda(i)$.
In parallel to results in Chapter \ref{chap:phase-transitions}, we show that several commonly used family-wise error rate-control procedures --- including Bonferroni's procedure --- are asymptotically optimal for the {exact}, and {exact-approximate} support recovery problems (as defined in Definition \ref{def:exact-recovery-success-failure}) in idealized chi-square models \eqref{eq:model-chisq-Chapter6} with independent components.
We further show that the \ac{BH} procedure is asymptotically optimal for the {approximate}, and {approximate-exact} support recovery problems.
% \ref{thm:chi-squared-exact-boundary}, \ref{thm:chi-squared-exact-approx-boundary}, \ref{thm:chi-squared-approx-boundary}, and \ref{thm:chi-squared-approx-exact-boundary}.
Under appropriate parametrizations of the signal sizes and sparsity, they establish the phase transitions of support recovery problems in the chi-square model.
Remarkably, the degree-of-freedom parameter does not affect the asymptotic boundaries in any of the four support recovery problems.

All phase transition boundaries coincide with those in the additive error models obtained in Chapter \ref{chap:phase-transitions} under suitable parametrizations.
% and Figure \ref{fig:phase-chi-squared} continues to apply.
indicating vanishing differences between the difficulties of the one-sided and two-sided alternatives in the Gaussian additive error model \eqref{eq:model-additive}.

\medskip

We then return to association screenings of categorical variables in
Section \ref{sec:odds-and-power}, and present the consequences of the phase transition in the exact-approximate problem in large-scale genetic association studies.
% present empirical evidence for the phase transition in the exact-approximate problem using real data from large-scale association studies on breast cancer obtained from the NHGRI-EBI GWAS Catalog \citep{macarthur2016new}.
% demystify the notion of signal size $\lambda$ in this context.
% which is perhaps less transparent than in additive error models.
We do so by characterizing the relationship between the signal size $\lambda$ and the marginal frequencies, odds ratio, and sample sizes for association tests on 2-by-2 contingency tables.
% Specifically, the amount of signal, when rare variants are present, is weaker compared to the signal when the marginal distributions are balanced.
% In other words, reliable detection of the effects by rare variants would require more samples compared to common variants, even at the same odds ratio.
This result, establishing the relationship between sample sizes and signal sizes, is made precise in Section \ref{sec:odds-and-power}.

We elaborate on the implications of this relationship on optimal study designs for association studies in Section \ref{sec:optimal-design}.
Perhaps surprisingly, our analysis reveals that balanced designs with equal number of cases and controls are often statistically inefficient.
Practical consequences of these results in power analysis will be illustrated with data examples in Section \ref{sec:phase-transitions-in-GWAS}.

The phase transitions in the chi-square models are demonstrated with numerical simulations in Section \ref{sec:numerical}.
Proofs of results in this Chapter are collected in Section \ref{sec:proof-signal-size-odds-ratio}.

% Practical issues are also addressed to make for simple and effective power analysis.

\section{Support recovery problems in chi-squared models}
\label{sec:chisq-boundaries}
\input{6.GWAS/6.chi-square.tex}

\section{Odds ratios and statistical power}
\label{sec:odds-and-power}
\input{6.GWAS/6.odds-and-power.tex}

\section{Optimal study designs and rare variants}
\label{sec:optimal-design} 
\input{6.GWAS/6.optimal-design.tex}

\section{Phase transitions in large-scale association screening studies}
\label{sec:phase-transitions-in-GWAS}
\input{6.GWAS/6.GWAS-phase-transitions.tex}

\section{Numerical illustrations of the phase transitions in chi-square models}
\label{sec:numerical}
\input{6.GWAS/6.numerical.tex}

\section{Proofs}
\label{sec:proof-signal-size-odds-ratio}
\input{6.GWAS/6.proofs-GWAS.tex}

\section{Exercises}
\label{sec:Exercises-GWAS}
\input{6.GWAS/6.Exercises.tex}