### 2020-08-21 acronyms.tex

\extrachap{Acronyms}

 \begin{description}[CABR]
  \item[\index{AGG}{AGG}]{asymptotically generalized Gaussian}
  \item[\index{BH}{BH}]{Benjamini-Hochberg}
 \item[\index{CDF}{CDF}]{cumulative distribution function}
 \item[\index{FDR}{FDR}]{false discovery rate}
 \item[\index{FNR}{FNR}]{false non-discovery rate}
 \item[\index{FWER}{FWER}]{family-wise error rate}
 \item[\index{FWNR}{FWNR}]{family-wise non-discovery rate}
 \item[\index{GWAS}{GWAS}]{genome-wide association studies}
 \item[\index{HC}{HC}]{higher-criticism}
 \item[\index{iid}{iid}]{independent and identically distributed}
 \item[\index{LR}{LR}]{likelihood-ratio}
 \item[\index{RS}{RS}]{relatively stable/relative stability}
 \item[\index{SNP}{SNP}]{single-nucleotide polymorphisms}
 \item[\index{URS}{URS}]{uniform relatively stable/uniform relative stability}
 \end{description}