### 2020-12-22 2.background/2.Exercises.tex
Addressed the minor typos of R2
I defined the command "\stilian" in the preamble and used it to make changes in response to the minor typos.

Please see reviews/GS_Annotated.pdf.

\stilian{R2, Typo 4}{I suggest we move all exercises to the online Supplement.  The supplement can have several chapters, the first of which could be 
"A crash course on extremes" or ``Selected background from EVT".  It could address some comments on the reviewers that the practitioners could need 
a lot of ``hand holding".  Similarly, in the Supplement we can put other ``basic" things we would like to have like ``Standard" derivations on the Sidak, BH procedure
the definition of Tukey's Higher Criticism statistics and the results from Empirical processes that are used to study it...} \fbox{ TO DO: Verify and add}

\begin{enumerate}

 \item (Inspired by \cite{resnick1973almost}) Let $\epsilon(i)\sim F,\ i=1,2,\cdots$ have arbitrary dependence 
 and $u_p = F^{\leftarrow}(1-1/p) \to \infty$.

 {\bf (a)}   If for some $\delta_p\to 0$,
 we have $\sum_{p} \overline{F}((1+\delta_p)u_p) <\infty$, then show that
 $$
 \limsup_{p\to\infty} \frac{M_p}{u_p} \le 1,\ \ \mbox{ almost surely,} 
 $$
 where $M_p:= \max_{i=1,\cdots,p} \epsilon(i)$.\\

 {\bf (b)} Show that the condition of part {\bf (a)} holds for the Generalized AGG$(\nu),\ \nu>0$ distributions.\\

 {\em Hint:} In part {\bf (a)}, argue that the events $\{ M_p > (1+\delta_p) u_p,\ \mbox{ infinitely often}\}$ and
  $\{\epsilon(p) > (1+\delta_p)u_p, \ \mbox{infinitely often}\}$ are equal. Appeal to the Borel Zero-One Law.\\

 \item  Show that independent realizations from a Generalized AGG$(\nu),\ \nu>0$ distribution
  are {\em almost surely stable}. That is, let $\epsilon(i)$'s be independent with distribution function $F$ as in Example \ref{exmp:AGG}.\\

{\bf (a)} Show that there exists a sequence $\delta_p\to 0$, such that 
$$
\sum_p \overline F((1+\delta_p)u_p) <\infty\ \ \mbox{ and }\ \  \sum_p F^p((1-\delta_p)u_p) <\infty.
$$
 {\bf (b)} By appealing to the Borel Zero-One law, conclude that in this case
 $$
 \frac{1}{u_p} M_p {\longrightarrow} 1,\ \ \mbox{ almost surely}.
 $$

{\em Comment:} 
\cite{resnick1973almost} provide a general necessary and sufficient condition for almost sure stability of independent maxima.\\

 \item Prove Gnedenko's result in part (2) of Proposition \ref{prop:rapid-varying-tails}.\\

\item Suppose that $F$ is heavy-tailed, i.e., $\overline F(x) \sim x^{-\alpha} L(x),$ as $x\to\infty$, for some $\alpha>0$ and a
slowly varying function $L$.  That is, $L(tx)/L(t)\to 1$, as $t\to\infty$, for all $x>0$.\\

{\bf (a)} Show that if $u_p$ is such that $p\overline F(u_p)\to 1$ as $p\to\infty$, then $p \overline F(x u_p)\to x^{-\alpha}$.\\

{\bf (b)} Let $\epsilon(i)$'s be independent realizations from $F$. With $u_n$ as in part {\bf (a)}, show that
$$ 
\frac{M_p}{u_p} \stackrel{d}{\longrightarrow} Z,\ \ \mbox{ as }p\to\infty,
$$ 
where $M_p = \max_{i=1,\cdots,p} \epsilon(i)$ and $\P(Z\le x) = e^{-1/x^\alpha},\ x>0$.\\

{\em Comment:} In the heavy-tailed case, rather than concentration of maxima (relative stability),
we encounter {\em dispersion of maxima} where under rescaling $M_p$ converges in distribution to a proper random variable.\\

\item Let  $M_p,\ p=1,2,\cdots$  be an infinite sequence of random variables such that
$$
\frac{M_p}{u_p} \stackrel{\P}{\longrightarrow} 1\ \ \mbox{ and }\ \ \frac{M_p}{v_p} \stackrel{d}{\longrightarrow} \xi,
$$
for some deterministic positive sequences  $u_p$ and $v_p$ and some non-zero random variable $\xi$.

Show that $\lim_{p\to\infty} u_p/v_p = c$ and $\xi = c$, almost surely.\\

{\em Hint:} Using Skorokhod's theorem there exist $\widetilde M_p$ and $\widetilde \xi$ on another probability space
such that $M_p/v_p\stackrel{d}{=}\widetilde M_p/v_p \stackrel{a.s.}{\to} \widetilde \xi \stackrel{d}{=}\xi$.
Then, recall the following criterion.

\begin{lemma} We have $\eta_p\stackrel{\P}{\to} \eta,\ p\to\infty$ if and only if, for every sequence $p_n\to \infty$, there is a further
subsequence $\{\widetilde p_n\}\subset\{p_n\}$, such that $\eta_{\widetilde p_n} \stackrel{a.s.}{\to} \eta$ almost surely, as $\widetilde p_n\to
\infty$.
\end{lemma}

\item Prove that the Benjamini-Hochberg procedure in Definition \ref{def:BH} controls FDR for independent data.\\

\stilian{R2, Typo 5}{ \fbox{TO DO} {\em Hint:} \fbox{ Give a good hint. } }

\end{enumerate}