### 2021-01-08 7.Supplement/supplement.tex

\chapter{Proofs for Chapter \ref{chap:phase-transitions}}
\label{sec:supplement:proofs}
\input{3.phase_transitions/3.proofs-phase-transitions.tex}

\chapter{Exact support recovery in non AGG models}

\section{Strong classification boundaries in other light-tailed error models}
\label{suppsec:other-boundaries}
\input{4.exact_support_recovery/4.other-boundaries.tex}

\section{Thresholding procedures under heavy-tailed errors}
\label{suppsec:heavy-tailed}
\input{4.exact_support_recovery/4.heavy-tailed.tex}