### 2021-01-13 4.exact_support_recovery/exact_support_recovery.tex
Chapter 4
Implementing the corrections...

We focus on exact support recovery problems in this chapter, and generalize the results we obtained in Chapter \ref{chap:phase-transitions} to additive error models with much relaxed distributional and dependence assumptions.

Consider the additive error model \eqref{eq:model-additive} with the triangular array of errors,
\begin{equation} \label{eq:error-array}
    {\cal E} = \left\{ (\epsilon_p(i))_{i=1}^p,\ p=1,2,\dots\right\},
\end{equation}
where the $\epsilon_p(i)$'s have common cumulative distribution function $F(x)=\P[\epsilon_p(i)\le x]$.
In contrast to the assumptions in Chapter \ref{chap:phase-transitions}, we only require the errors to have common marginal distributions, and allow them to have potentially arbitrary dependence.
% We will study the role of dependence in support recovery problems in Section \ref{sec:URS}.

Although our method of analysis applies to all light-tailed error distributions with rapidly varying tails (see Definition \ref{def:rapid-variation}), to be concrete and better convey the main ideas, we will focus on the class of $\mathrm{AGG}(\nu)$ laws (see Definition \ref{def:AGG}).
\stilian{Extensions of the results to other error models are presented in Section \ref{suppsec:other-boundaries}.}{
\fbox{But, we may bring them back if eliminating the proofs from Ch. 6 gives us enough space!}}

As in Chapter \ref{chap:phase-transitions}, we assume the signal in model \eqref{eq:model-additive} to be a sparse vector $\mu_p = \left(\mu_p(i)\right)_{i=1}^p$, with support
$
S_p:= \{ i\, :\, \mu_p(i)\not = 0\}.
$ 
The signal sparsity, with a few exceptions which will be explicitly stated, is parametrized as
\begin{equation} \label{eq:sparsity-parametrized}
    s = |S_p| = \lfloor p^{1 -\beta} \rfloor, % \quad \text{with } 0 < \beta \le 1 \; \text{(fixed)},
\end{equation}
with $0 < \beta \le 1$ fixed.

We assume that the non-zero entries of $\mu$ are positive and take values in the interval $\left[\underline{\Delta},\overline{\Delta}\right)\subset (0,\infty)$.
That is, $0<\underline{\Delta}\le\mu(i)<\overline{\Delta}\le+\infty$, for all $i\in S_p$.
The lower and upper bound on the signal sizes $\underline{\Delta}$ and $\overline{\Delta}$ are parametrized as
\begin{equation} \label{eq:signal-size-parametrized}
    \underline{\Delta} = \underline{\Delta}(p) = (\nu \underline{r} \log{p})^{1/\nu} \quad \text{and} \quad
    \overline{\Delta} = \overline{\Delta}(p)  = (\nu \overline{r} \log{p})^{1/\nu},
\end{equation}
with parameters $0 < \underline{r} \le \overline{r} \le +\infty$.
Notice that the parametrization now depends on the shape of the assumed error distributions $\mathrm{AGG}(\nu)$ through the parameter $\nu>0$.

According to Lemma \ref{lemma:risk-exact-recovery-probability}, in order to study the asymptotic behaviors of 
the {\em exact support recovery} $\mathrm{risk}^{\mathrm{E}}$, it is sufficient to establish conditions under which the support sets can be consistently estimated, i.e.,
\begin{equation} \label{eq:exact-recovery} 
    \P[\widehat{S}_p = S_p] \longrightarrow 1 \quad \text{as } \; p\to\infty, 
\end{equation}
where $\widehat{S}_p$ is an estimate of the true signal support set $S_p$.

\medskip

Several authors have studied the support recovery problem in terms of the Hamming loss and obtained minimax optimality results \citep[see, e.g.][]{ji2012ups, genovese2012comparison, jin2014optimality, butucea2018variable}.
% We briefly review the recent work by \cite{butucea2018variable}, whose .
In the special case of Gaussian marginals, \cite{butucea2018variable} showed that the boundary \eqref{eq:strong-classification-boundary-Gaussian} exists in a minimax sense.
That is, when errors are \emph{independent} Gaussians and the signal sizes fall below the boundary 
\eqref{eq:strong-classification-boundary-Gaussian}, then no procedure can have a vanishing Hamming
loss, as $p\to\infty$.  Conversely, if signal sizes are above that same boundary, the Hamming loss can be made to 
vanish with a specific thresholding procedure.

Unfortunately, as pointed out in Section \ref{sec:risks-relations}, vanishing Hamming loss is only sufficient, but
not necessary for support recovery \eqref{eq:exact-recovery}, and the sharp results of \cite{butucea2018variable}, 
do not carry over directly to the study of the probability of support recovery or the exact support recovery risk.
More importantly, despite their elegance, these Hamming loss-minimax studies naturally reduce to the analysis of the special case of iid errors, which is by design blind to non-trivial error-dependence structures.
%%This prevents us from fully exploring of the phase transition phenomena under other dependence conditions.
% This is arguably a consequence of the inherent limitation of the Hamming-loss approach. 
% For practitioners, a minimax statement in the Gaussian case seems to offer little guidance.
% Indeed, in applications, independence is an exception rather than the rule; Gaussianity of the errors may also be unnecessarily restrictive.

So far in the literature, the role of dependence, and that of the distributional assumptions in model \eqref{eq:model-additive} remained largely unexplored.  
This chapter offers advances in both directions, and provides a close-to-complete solution of the exact support recovery problem. (See also, Chapter \ref{chap:URS}.)
We briefly summarize our contributions next.

\medskip

In Section \ref{sec:boundary}, we study exact support recovery in the sense of \eqref{eq:exact-recovery} directly, under general distributional and dependence assumptions. 
In particular, we describe the phase transition phenomena in the dependent AGG model, under the scaling described in \eqref{eq:sparsity-parametrized} and \eqref{eq:signal-size-parametrized}.
Consider the function
\begin{equation} \label{eq:strong-classification-boundary}
    g(\beta) = g_\nu(\beta) = (1 + (1 - \beta)^{1/\nu})^\nu, \quad \nu>0,
\end{equation}
which we refer to as the {\em strong classification boundary}.
In Theorem \ref{thm:sufficient} we show that, 
% under the general distributional assumptions in Section \ref{subsec:set-up}, 
if the signal sizes are above the boundary (i.e., $\underline{r}> g(\beta)$), the \ac{FWER}-controlling procedures described in Section \ref{sec:statistical-procedures} with appropriately calibrated levels achieve \emph{exact support recovery} as in \eqref{eq:exact-recovery}.

Conversely, under mild error-dependence conditions, we show in Theorem \ref{thm:necessary}, that
no thresholding procedure can achieve the asymptotically perfect support recovery \eqref{eq:exact-recovery}
when the signal size is below the boundary  (i.e., $r<g(\beta)$).   In fact, in this case, we have
\begin{equation} \label{eq:exact-recovery-failure}
    \mathbb{P}\left[\widehat{S}_p=S_p\right]\longrightarrow 0,\quad \mbox{ as }p\to \infty,
\end{equation}
for all $\widehat{S}_p$ in the form of \eqref{eq:thresholding-procedure}.  This strong 
impossibility of exact support recovery result holds for a surprisingly large class of dependence structures 
characterized by the concept of \emph{uniform relative stability} (URS, see Definition \ref{def:URS} below).

These two results show that the thresholding procedures obey a phase transition phenomenon in a strong, 
\emph{point-wise} sense over the class of errors with AGG$(\nu),\ \nu>0$ marginal distributions {\em and} URS 
dependence. The conclusions are fundamentally stronger and more informative than the minimax statements in the literature \citep[see, e.g.,][]{butucea2018variable}.
The techniques developed here are also entirely different from those in \citet{ji2012ups} or \citet{butucea2018variable}.
\stilian{It should be noted that the URS condition is fundamentally weaker than the conventional ``mixing conditions'' employed in the analysis of sums and averages.  This is because the key to exact support recovery is the study of extremes
(viz maxima) which have fundamentally different behavior than sums, under dependence.  More insights and transparent characterizations of the dependence conditions under which the phase transition type result holds will be established in the Chapter \ref{chap:URS}.}{\fbox{check}}

% The strong classification boundary $g$ characterizes a phase-transition phenomenon similar to that of the signal detection and approximate support recovery. 
% A preview of this result is presented in Figure \ref{fig:phase}.

% \begin{figure}
%       \centering
%       \includegraphics[width=0.45\textwidth]{./figures/phase_diagram_Gaussian_no_dashed_area.eps}
%       \includegraphics[width=0.45\textwidth]{./figures/phase_diagram_double_exponential.eps}
%       \caption{The phase diagrams of the detection, weak classification, and strong classification boundaries against sparse alternatives under Gaussian (left) and Laplace distributed (right) errors. Here $\beta$ and ${r}$ parametrize the signal sparsity and the lower and upper bounds of the signal sizes, respectively. 
%       The signal detection problem can be answered perfectly (asymptotically) inside the \emph{Detectable} region; false discovery proportion and non-discovery proportion can be made to vanish in the \emph{Weakly classifiable} region. We study in this paper the strong classification boundary, above which the support recovery can be achieved \emph{exactly} in the \emph{Classifiable} region $\{(\beta, {r}):{r}>g(\beta)\}$. 
%       In a large class of dependence structures characterized by URS, when signal sizes fall below the strong classification boundary \eqref{eq:strong-classification-boundary}, i.e. $\{(\beta, {r}):{r}<g(\beta)\}$, no thresholding procedure succeeds in the exact support recovery problem. }
%       \label{fig:phase}
% \end{figure}

\medskip

We show in Section \ref{sec:optimality} that data thresholding procedures are not only asymptotically, but in fact, finite sample optimal when the errors are independent and identically distributed (iid) with log-concave densities. 
In this case, no estimator can achieve perfect support recovery when the signal is below the strong classification boundary \eqref{eq:strong-classification-boundary}.  
Consequently, in the case of $\text{AGG}(\nu)$ errors with $\nu\ge 1$, the strong 
classification boundary is shown to hold in the minimax 
sense for \emph{all} procedures. This is formalized in Theorem \ref{thm:necessary-strengthened} and Corollary \ref{cor:minimax-over-dependence}.

\medskip
The phase transition phenomena for two additional classes of error distributions with either heavier or lighter tails than the AGG distributions will be described in Section \ref{suppsec:other-boundaries}.

\medskip

A final surprising result that had only recently been noticed by the statistical community is that thresholding procedures, including data-dependent ones, are not optimal in general in the support recovery problem when the errors have heavy (regularly-varying) tails. 
\citet{arias2019detection} discussed the phenomena in approximate support recovery problems.
In this case, we also demonstrate the absence of a phase transition phenomenon in exact support recovery by thresholding, in Section \ref{suppsec:heavy-tailed}.

% In this sense, for $\nu\ge 1$ our results are rather complete.  We note here that our focus is not on minimax analysis of the support recovery problem, where a large class of distributional and dependence structures are considered.  
% In some sense, the worst case scenario (see also \cite{butucea2018variable}) is when the errors are independent.  
% In contrast, we establish the phase-transition phenomenon, for fixed but very general dependence conditions characterized by the URS property.

%We show that when the error-distribution $F$ is log-concave, then the thresholding procedures are optimal (for independent errors).  Therefore, in this regime the {\em strong classification boundary} $g$ is universal.  
%That is, if $r<g(\beta)$, no estimator can achieve perfect support recovery as $p\to\infty$. 
% \fbox{maybe skip the next sentence} 
% In fact, going beyond the class of rapidly varying distributions, (e.g., for heavy Pareto-type tails) the support recovery problem is fundamentally different and there may no longer be a phase-transition phenomenon.

\section{Exact support recovery under AGG errors}
\label{sec:boundary}
\input{4.exact_support_recovery/4.phase_transition_results.tex}

\section{Bayes minimax optimality and (sub)optimality of thresholding procedures}
\label{sec:optimality}
\input{4.exact_support_recovery/4.minimax.tex}

%\section{Strong classification boundaries in other light-tailed error models}
%\label{suppsec:other-boundaries}
%\input{4.exact_support_recovery/4.other-boundaries.tex}

%\section{Thresholding procedures under heavy-tailed errors}
%\label{suppsec:heavy-tailed}
%\input{4.exact_support_recovery/4.heavy-tailed.tex}

\section{Additional proofs}
\label{suppsec:proofs}
\input{4.exact_support_recovery/4.proofs-examples.tex}
\input{4.exact_support_recovery/4.proofs-convexity.tex}