### 2021-01-15 4.exact_support_recovery/4.pointwise.minimax.tex
.

We establish in this section minimax versions of our results from Section \ref{sec:boundary}.
Specifically, if we restrict ourselves to \emph{the class of thresholding procedures} $\cal T$ (defined in \eqref{eq:thresholding-procedure}), then Bonferroni's procedure is minimax optimal, for \emph{any} fixed dependence structures in the URS class.
This is formalized in Corollary \ref{cor:point-wise-minimax} in Section \ref{subsec:point-wise-minimax}.
We refer to this result as \emph{point-wise} minimax, to emphasize the fact that this optimality holds for every \emph{fixed} URS array.

Meanwhile, if we search over \emph{all procedures}, but expand the space of models to include {\em all} dependence structures, then a different minimax optimality statement holds for Bonferroni's procedure.
This result, formally stated in Section \ref{subsec:minimax-over-dependence}, is a consequence of our characterization of the finite-sample Bayes optimality of thresholding procedures in Sections \ref{subsec:Bayes-optimality} and \ref{subsec:optimal-procedure-sub-exponential}.
% The result is of independent interest and it is the main contribution of this section.

Finally, we offer some insights into the support recovery problem in the case when errors have heavier-than-exponential tails in Section \ref{subsec:optimal-procedure-super-exponential}.

\subsection{Point-wise minimax optimality}
\label{subsec:point-wise-minimax}

Theorems \ref{thm:sufficient} and \ref{thm:necessary} can be cast in the form of an asymptotic minimax statement. 
% (concealing the gap between sufficient and necessary conditions as pointed out in Remark \ref{rmk:gap-between-sufficient-necessary}).

\begin{corollary}[Point-wise minimax]
\label{cor:point-wise-minimax}
Let $\widehat{S}^{\text{Bonf}}$ be the sequence of Bonferroni's procedure described in Theorem \ref{thm:sufficient}. 
Let also the errors have common $\text{AGG}(\nu)$ distribution $F$ with parameter $\nu>0$, and $\Theta_p^+$ be as defined in \eqref{eq:minimax-signal-config-over}.
If $\underline{r}>g(\beta)$, then we have
\begin{equation} \label{eq:point-wise-minimax-above}
    \limsup_{p\to\infty} \sup_{\mu\in\Theta_p^+(\beta, \underline{r})} \P(\widehat{S}^{\text{Bonf}}_p \neq S_p) = 0,
\end{equation}
for arbitrary dependence structure of the error array ${\cal E} = \{\epsilon_p(i)\}_p$.
Let $\cal T$ be the class of thresholding procedures \eqref{eq:thresholding-procedure}. 
If $\underline{r}<g(\beta)$, then we have
\begin{equation} \label{eq:point-wise-minimax-below}
    \liminf_{p\to\infty} \inf_{\widehat{S}_p\in \cal T} \sup_{\mu\in\Theta_p^+(\beta, \underline{r})} \P(\widehat{S}_p \neq S_p) = 1,
\end{equation}
for any error dependence structure such that ${\cal E}\in U(F)$ and ${(-\cal E)}\in U(F)$.
\end{corollary}

\begin{proof}[Proof of Corollary \ref{cor:point-wise-minimax}]
The first conclusion \eqref{eq:point-wise-minimax-above} is a restatement of Theorem \ref{thm:sufficient}.

For the second statement \eqref{eq:point-wise-minimax-below}, since $\underline{r}<g(\beta)$, we can pick a sequence $\mu^*\in\Theta_p^+(\beta, \underline{r})$ such that $|S_p| = \lfloor p^{1-\beta}\rfloor$, with signals having the same signal size $\mu(i)=(2r\log{p})^{1/\nu}$ for all $i\in S_p$, where $\underline{r}<{r}<g(\beta)$.
For this particular choice of $\mu^*$ we have $\mu^*\in\Theta_p^-(\beta, \overline{r})$ where $r<\overline{r}<g(\beta)$,
and according to Theorem \ref{thm:necessary}, we obtain $\lim_{p\to\infty} \inf_{\widehat{S}_p\in \cal T} \mathbb P[\widehat{S}_p \neq S_p] = 1$, for all dependence structures in the URS class.
\end{proof}

\begin{remark}
Theorem \ref{thm:necessary} is a stronger result than the traditional minimax claim in Relation \eqref{eq:point-wise-minimax-below}.
Indeed,  \eqref{eq:classification-impossible-dependent} involves an infimum (over the class $\Theta^-_p$) while \eqref{eq:point-wise-minimax-below} has a supremum (over the class $\Theta^+_p$).

On the other hand, Corollary \ref{cor:point-wise-minimax} is more informative than many minimax-type statements, since it applies ``point-wise'' to any fixed error dependence structure in the URS class.
\end{remark}

\begin{remark}
Corollary \ref{cor:point-wise-minimax} echoes Remark \ref{rmk:dependence-assumptions}: for a very large class of dependence structures, we cannot improve upon Bonferroni's procedure in exact support recovery problems (asymptotically), unless we look beyond thresholding procedures.
% There is a limited amount of literature on non-thresholding procedures that utilize the error dependence structures to improve power.
% A notable effort was made by \citet{jin2014optimality}, where the performance (in terms of Hamming loss) of a so-called graphlet screening procedure was studied. 
\end{remark}

% This is an enhancement (in the asymptotic sense) of the minimax results in Section 4.1 of \cite{butucea2018variable}, where the supremum was taken over the dependence structures.