### 2021-01-16 4.exact_support_recovery/4.proofs-examples.tex
update
cleaned up Ch 4.

Both (new) Ch 4 and 5 are ready to re-read.

\subsection{Proof of the claims in Examples \ref{exmp:FWER-controlling_procedures} and \ref{exmp:signals-straddling-the-boundary}}
\label{subsec:proofs-examples}