### 2021-01-29 3.phase_transitions/3.proofs-phase-transitions.tex
ch 3
almost done editing ch 3 (for formatting).  I need to absorb the auxuliary results from the appendix and check the proofs carefully once again. I like the new format where the proofs are presented right after the statments and the various support recovery criteria are presented in separate sections rather than sub-sections.  Thus, they are visible in the table of contents.

We first recall some basic properties of the Gaussian distribution in Section \ref{sec:Gaussian-distributions}.
Section \ref{suppsec:BH-monotonicity} states and proves an interesting property of the \ac{BH} procedure which may be of independent interest.
Results on the signal detection problem (Theorem \ref{thm:detection-optimality}) are proved in Section \ref{subsec:proof-additive-error-detection-boundaries}, and the phase transition results on the support recovery problems (Theorems \ref{thm:Gaussian-error-exact-boundary} through \ref{thm:Gaussian-error-approx-exact-boundary}) are shown in Sections \ref{subsec:proof-additive-error-approx-boundaries} and \ref{subsec:proof-additive-error-mix-boundaries}.

\section{Auxiliary facts of Gaussian distributions}
\label{sec:Gaussian-distributions}
\input{3.phase_transitions/3.Gaussian-distributions.tex}

\section{Proof of Theorem \ref{thm:detection-optimality}}
\label{subsec:proof-additive-error-detection-boundaries}

\section{Proof of Theorem \ref{thm:Gaussian-error-approx-boundary}}
\label{subsec:proof-additive-error-approx-boundaries}

\section{Proof of Theorems \ref{thm:Gaussian-error-exact-approx-boundary} and \ref{thm:Gaussian-error-approx-exact-boundary}}
\label{subsec:proof-additive-error-mix-boundaries}