### 2021-02-01 3.phase_transitions/phase_transitions.tex

The purpose of this chapter is to provide a unified review of the fundamental statistical limits in the 
sparse signal detection and support recovery problems.  Our goal is to convey the main ideas and thus we shall focus on 
the simple but important setting of independent Gaussian errors.
Specifically, we derive the conditions under which the detection and support recovery problems succeed and fail in the sense of \eqref{eq:support-recovery-success} and \eqref{eq:support-recovery-failure}, in the additive error model 
\begin{equation} \label{eq:model-additive-Chapter3}
    x(i) = \mu(i) + \epsilon(i), \quad i=1,\ldots,p,
\end{equation}
where the errors $\epsilon(i)$'s are \ac{iid} standard Gaussians random variables.
Once again, we restrict our analysis to models with independent and identically distributed Gaussian errors for the 
moment. Both the distributional assumption and the independence assumption will be relaxed substantially in the following chapters.

As laid out in Section \ref{sec:asymptotics}, we work under the asymptotic regime where the problem dimension 
$p$ diverges to infinity.  The set of non-zero entires of the signal vector $\mu = \mu_p$ will be referred
to as its {\em support} and denoted by 
$$
 S_p:=\{i\, :\, \mu(i)\not = 0\}.
$$
We shall assume that the size of the support is
\begin{equation} \label{eq:signal-sparsity-additive}
    |S_p| = \left\lfloor p^{1-\beta} \right\rfloor, \quad \beta\in(0,1],
\end{equation}
where $\beta$ parametrizes the problem sparsity.
The closer $\beta$ is to 1, the sparser the support $S_p$.  
Conversely, when $\beta$ is close to 0, the support is dense with many non-null signals.
We consider one-sided alternatives \eqref{eq:global-test-one-sided}, and parametrize the range of the non-zero (and perhaps unequal) signals with
\begin{equation} \label{eq:signal-size-additive}
    \underline{\Delta} = \sqrt{2\underline{r}\log{p}}
    \le \mu(i) \le
    \overline{\Delta} = \sqrt{2\overline{r}\log{p}}, \quad \text{for all}\;\;i\in S_p,
\end{equation}
for some constants $0<\underline{r}\le\overline{r}\le+\infty$.

% \label{rmk:global-test-boundary}
The parametrization of signal sparsity \eqref{eq:signal-sparsity-additive} and signal sizes \eqref{eq:signal-size-additive} in the Gaussian model was first introduced in \citet{ingster1998minimax}, and later adopted by \cite{hall2010innovated}, \cite{cai2011optimal}, \cite{zhong2013tests}, \cite{cai2014optimal}, \cite{arias2017distribution1}, and numerous others for studying the signal detection problem in Gaussian location-scale models.
Similar scalings of sparsity and signal size are also used in, e.g., \cite{ji2012ups}, \cite{jin2014optimality}, \cite{butucea2018variable} to study the phase transitions of the support recovery problems under Gaussianity assumptions.

\section{Sparse signal detection problems}
\label{sec:global-tests}
\input{3.phase_transitions/3.global-tests.tex}

%\newpage
\section{Sparse signal support recovery problems}
\label{sec:additive-error-model-boundaries}
\input{3.phase_transitions/3.Gaussian-support-recovery.tex}

%\section{Proofs}
%\label{sec:proofs}
%\input{3.phase_transitions/3.proofs-phase-transitions.tex}