### 2021-02-01 2.background/2.Gaussian-distributions.tex
all in
I completed the merge of my work on ch. 3, 4 and 5 with the branch "revision-GWAS".  Ready for a "pull" to "revision-GWAS"

We end this chapter with several facts about univariate and multivariate Gaussian distributions that will be used in the rest of the manuscript. \\

{\bf Relative stability.} We first state the relative stability of iid standard Gaussian random variables.  Since the standard Gaussian distribution falls in the 
class of asymptotically generalized Gaussians (AGG; see Definition \ref{def:AGG}), by Example \ref{exmp:AGG}, we know that the triangular array ${\cal E} = \{\left(\epsilon_p(i)\right)_{i=1}^p, p\in\N\}$ has relatively stable (RS) maxima in the sense of \eqref{eq:RS-condition}, i.e.,
\begin{equation} \label{eq:relative-stability-Gaussian-maxima}
    \frac{1}{u_{p}} \max_{i=1,\ldots,p} \epsilon_p(i) \xrightarrow{\P} 1,\quad \text{as }\;p\to\infty,
\end{equation}
where $u_p$ is the $(1/p)$-th upper quantile as defined in \eqref{eq:AGG-quantiles}.
Similarly, since the array ${\cal E}$ has distributions symmetric around 0, it also has relatively stable minima
\begin{equation} \label{eq:relative-stability-Gaussian-minima}
    \frac{1}{u_{p}} \min_{i=1,\ldots,p} \epsilon_p(i) \xrightarrow{\P} -1,\quad \text{as }\;p\to\infty.
\end{equation}
The convergence in \eqref{eq:relative-stability-Gaussian-maxima} also holds almost surely.\\

{\bf Mill's ratio.} We give next the well-known bounds for the Mill's ratio of Gaussian tails.
Let $\Phi$ denote the CDF of the standard Gaussian distribution and $\phi$ its density.
One can show that for all $x>0$ we have
\begin{equation} \label{eq:Mills-ratio}
    \frac{x}{1+x^2}\phi(x) \le \overline{\Phi}(x) = 1-\Phi(x) \le \frac{1}{x}\phi(x),
\end{equation}
using e.g., integration by parts.  Note that this fact may be used to verify the rapid variation of $\Phi$, which entails 
the relative stability property above.\\

{\bf Stochastic monotonicity.} The third fact is the stochastic monotonicity of the Gaussian location family. 
In fact, for all location families $\{F_\delta(x)\}_\delta$ where $F_\delta(x) = F(x-\delta)$, we have,
\begin{equation} \label{eq:stochastic-monotonicity-Gaussian}
    F_{\delta_1}(t) \ge F_{\delta_2}(t), \quad \text{for all}\quad t\in\mathbb{R}\quad\text{and all}\quad \delta_1 \le \delta_2.
\end{equation}
Relation \eqref{eq:stochastic-monotonicity-Gaussian} holds, of course, when $F$ is the standard Gaussian distribution. \\

{\bf Slepian's lemma and the Sudakov-Fernique inequality.} 
The following two result will play fundamental roles in our characterization of uniform relative stability for Gaussian triangular arrays in 
Chapter \ref{chap:URS}. 
% Many more details and deeper insights can be found for example in the monographs 
%\cite{leadbetter2012extremes} and \cite{adler2009random}, among others. 
The first is the celebrated result due to \citet{slepian1962one}.

\begin{theorem}[Slepian's Lemma]\label{thm:Slepian-lemma} Let $\epsilon = (\epsilon(i))_{i=1}^p$ and $\eta = (\eta(i))_{i=1}^p$ be 
two multivariate normally distributed random vectors with zero means $\E [\epsilon(i)] = \E[\eta(i)] = 0$.

 If for all $i,j=1,\cdots,p$, we have
$$
\E [\epsilon(i)^2] = \E[\eta(i)^2],\ \ \ \mbox{ and }\ \ \ {\rm Cov}(\epsilon(i),\epsilon(j)) \le {\rm Cov}(\eta(i),\ \eta(j)),
$$
 then $\epsilon \stackrel{{\rm st}}{\ge} \eta$, i.e.,
$$
\P [ \epsilon(i) \le x_i,\ i=1,\cdots,p] \le \P [ \eta(i) \le x_i,\ i=1,\cdots,p].
$$
\end{theorem}

This result implies in particular that $M_\epsilon:= \max_{i=1,\cdots,p}\epsilon(i)$ dominates stochastically $M_\eta:= \max_{i=1,\cdots,p} \eta(i)$.
Thus, for example, the maximum of iid Gaussians are stochastically larger the maxima any positively correlated Gaussian vector 
with the same marginal distributions.

Slepian's lemma can be obtained as a corollary from the general Normal Comparison Lemma 
\citep[see, e.g., Theorem 4.2.1 on page 81 in][]{leadbetter2012extremes}.  See also Ch.\ 2 in \cite{adler2009random}.

The following result, known as the Sudakov-Fernique inequality is similar in spirit to Slepian's lemma but does not assume that the Gaussian 
vectors are centered. For a proof and, in fact, a more general result, see e.g., Theorem 2.2.5 on page 61 in \cite{adler2009random}.

\begin{theorem} [Sudakov-Fernique Inequality] \label{th:Sudakov-Fernique} 
Let $\epsilon = (\epsilon(i))_{i=1}^p$ and $\eta = (\eta(i))_{i=1}^p$ be 
two multivariate normally distributed random vectors.

If for all $i,j=1,\cdots,p$, we have 
$$
\E [ \epsilon(i) ] = \E [\eta(i)] \ \ \mbox{ and }\ \ \E [ (\epsilon(i)-\epsilon(j))^2] \le \E [ (\eta(i)-\eta(j))^2],
$$
then for $M_\epsilon = \max_{i=1,\cdots,p}\epsilon(i)$ and $M_\eta = \max_{i=1,\cdots,p} \eta(i)$, we have 
$$
 \E [ M_\epsilon ] \le \E [M_\eta].
$$
\end{theorem}