### 2021-02-03 5.Bayes.optimality/5.Bayes_optimality_and_minimax.tex
DONE!
Ready for one final pass. Specifically, in the PDF, pls look at pages:

29, 30, 60 (notation <...>) Proofs of Propositons 5.1 and 5.3, Def. 5.3.

Page 115.

There is only one UNDEFINED reference on page 115. I can fix it once I understand the context better.

In this chapter, we investigate the universality of the phase-transition results on exact support recovery established in
Chapter \ref{chap:exact-support-recovery}.  Specifically, we would like to know to what extent the signal classification 
boundary valid for thresholding estimators applies to the best possible support estimators.  The answer to this question 
will complete the characterization of the fundamental limits in exact support recovery.

To this end, we begin by characterizing the finite-sample Bayes optimality of the thresholding procedures (Section
\ref{subsec:optimal-procedure-sub-exponential}). As we will see, the so-called oracle thresholding estimators are in 
fact finite-sample optimal for many classes of models.  These optimality results allow us to establish 
powerful minimax formulations of the exact support recovery phase-transition phenomenon showing its universality for all estimators (Theorem \ref{thm:necessary-strengthened}).

Perhaps surprisingly, however, thresholding estimators can be sub-optimal.  This is so for example 
in the additive noise model with heavier than exponential error densities (Example 
\ref{exmp:suboptimal-data-thresholding}).   In this case, {\em likelihood ratio
thresholding} rather than {\em data thresholding} can lead to optimal support estimators
(Section \ref{subsec:optimal-procedure-super-exponential}).  In such cases, the universality 
of the exact support recovery boundary remains an open question.

\section{Bayes optimality in support recovery problems}
\label{subsec:Bayes-optimality}

In studying support recovery problems \citep[e.g.,][]{arias2017distribution}, restrictions to the thresholding procedures 
are sometimes justified by arguing that such procedures are the ``reasonable'' choice for estimating the support set.
We show in this chapter that, perhaps surprisingly, for general error models, thresholding procedures are not always 
optimal, even when the observations are independent.

We shall identify the optimal procedure for support recovery problems under a Bayesian setting with general distributional assumptions (including but not limited to additive models \eqref{eq:model-additive}).
Specifically, we assume that there is an ordered set $P = (i_1,\ldots,i_s)$, $i_i\in\{1,\ldots,p\}$, and $s$ (not necessarily equal) densities $f_{1}, \ldots, f_{s}$, such that the observations indexed by $P$ have corresponding densities. That is,
\begin{equation} \label{eq:signal-distributions-ordered}
x(i_j) \sim f_j, \quad j=1,\ldots,s.
\end{equation}
Let also the rest $(p-s)$ observations have common density $f_0$, i.e., $x(i)\sim f_0$ for $i\not\in S$.
We further assume that the observations $x$ are mutually independent.

We adopt here a Bayesian framework to measure statistical risks. 
Let the ordered support $P=(i_1,\ldots,i_s)$ have prior
\begin{equation} \label{eq:uniform-ordered}
\pi((i_1,\ldots, i_s)) = {(p-s)!}/{p!},
\end{equation}
for all distinct $1\le i_1, \ldots, i_s\le p$.
Consequently, the unordered support $S=\{i_1,\ldots,i_s\}$ is distributed uniformly in the collection of all set of size $s$, with the unordered uniform distribution $\pi^{\text{u}}$. That is, for all for all $S\in\mathcal{S}:=\left\{S\subseteq\{1,\ldots,p\};|S|=s\right\}$, we have 
\begin{equation} \label{eq:uniform}
\pi^{\text{u}}
(\{i_1,\ldots, i_s\}) = \sum_{\sigma}\pi((i_{\sigma(1)},\ldots, i_{\sigma(s)})) = {(p-s)!s!}/{p!},
\end{equation}
where the sum is taken over all permuations of $\{1,2,\ldots,s\}$.

For any fixed configuration $P$, consider the loss function,
$$
\ell(\widehat{S},S) := \P[\widehat{S} \neq S] = \P_{P}[\widehat{S} \neq S],
$$
where the probability is taken over the randomness in the observations only.
The Bayes optimal procedures should minimize 
\begin{equation} \label{eq:Bayes-risk}
    \E_\pi \P[\widehat{S} \neq S],
\end{equation}
where the expectation is taken over the random configurations $P$, with a uniform distribution $\pi$ as specified in \eqref{eq:uniform-ordered}.

% For a sequence of estimators $\widehat{S} = \widehat{S}_p$, the asymptotically optimal procedures are defined to be the ones that minimize the asymptotic Bayes risk,
% \begin{equation} \label{eq:Bayes-risk}
%     R(\widehat{S}) := \limsup_{p\to\infty} \E_{P,{\cal E}} [\ell(\widehat{S},S)] = \limsup_{p\to\infty} \P_{P,{\cal E}}[\widehat{S} \neq S].
% \end{equation}

If, however, the sparsity $s = |S|$ of the problem is known, then a natural estimator for $S$ would be based on the set of top $s$ order statistics. Such estimators will be referred to as oracle thresholding estimators and formally defined next.

%\stilian{ 
For any collection of numbers $\{a_i,\ i=1,\cdots,s\}$, let 
$$
\langle a_1,\cdots,a_s\rangle := (a_{[1]},\cdots,a_{[s]})
$$
denote the vector of $a_i$'s arranged in a non-increasing order.

\begin{definition}[Oracle data thresholding] Let $x_{[1]} \ge \ldots \ge x_{[p]}$ be the order statistics of
the data vector $x$.  Any estimator $\widehat{S}^* :=\{ i_1,\cdots,i_s\}$, where
 $$
  \langle x(i_1),\cdots,x(i_s)\rangle  = (x_{[1]},\cdots,x_{[s]})
 $$ 
 will be referred to as an {\em oracle thresholding estimator}.
 \end{definition}

 Simply put, the oracle thresholding estimators are comprised of the indices corresponding to the $s$ largest values in the
 data.  Note that, in the absence of ties among the largest $s+1$ data values, the oracle thresholding estimator is unique.  
 For concreteness,  one can break possible ties lexicographically.  In many cases, 
 the oracle thresholding estimators will be almost surely unique.
 %}
%{\fbox{IMPORTANT, replacing the following}
%\begin{definition}[Oracle data thresholding]
%We call $\widehat{S}^* = \{i\,|\, x(i)\ge x_{[s]}\}$ the oracle data thresholding procedure, where $x_{[1]} \ge \ldots \ge x_{[p]}$ are the order statistics of $x$.
%\end{definition}}

\section{Bayes optimality of oracle thresholding} % under sub-exponential errors}
\label{subsec:optimal-procedure-sub-exponential}

In this section, we study the Bayes optimality of the oracle thresholding procedures.  
The following \emph{monotone likelihood ratio} (MLR) property will play a key role.

\begin{definition}[Monotone Likelihood Ratio]
A family of positive densities on $\R$, $\{f_\delta, \delta \in U\}$, is said to have the MLR property if, for all $\delta_0, \delta_1\in U\subseteq\R$ such that $\delta_0 < \delta_1$, the likelihood ratio $\left(f_{\delta_1}(x)/f_{\delta_0}(x)\right)$ is an increasing function of $x$.
\end{definition}

The next result provides a general criterion for the finite-sample Bayes optimality of the oracle thresholding 
procedure $\widehat{S}^*$.

\begin{proposition} \label{prop:optimal-oracle-procedures}
Let the observations $x(i)$, $i=1,\ldots,p$ be as prescribed as in \eqref{eq:signal-distributions-ordered} through \eqref{eq:uniform-ordered}.
If each of the pairs $\{f_0, f_{1}\},\ldots,\{f_0,f_{s}\}$ forms an MLR family, then every oracle data thresholding procedure $\widehat{S}^*$ is finite-sample optimal in terms of Bayes risk $\E_\pi \P[\widehat{S} \neq S]$. That is,
\begin{equation} \label{eq:finite-sample-Bayes-optimal}
    \widehat{S}^* \in \argmin_{\widehat{S}} \E_\pi \P[\widehat{S} \neq S].
\end{equation}
for all $s$ and $p$.
\end{proposition} 
\begin{proof}%[Proof of Proposition \ref{prop:optimal-oracle-procedures}]
The problem of support recovery can be equivalently stated as a classification problem, where the discrete parameter space is $\mathcal{S} = \{S\subseteq\{1,\ldots,p\}:|S|=s\}$, and the observation $x \in\R^p$ has likelihood $f(x|S)$ indexed by the support set $S$.

By the optimality of the Bayes classifier \citep[see, e.g.,][]{domingos1997optimality}, a set estimator that maximizes the probability of support recovery is such that
$$
\widehat{S} \in \argmax_{S\in \mathcal{S}} f(x|S) \pi(S).
$$
Since we know from \eqref{eq:uniform} that $\pi(\cdot)$ is uniform, the problem in our context reduces to showing that $f(x|\widehat{S}^*) = f(x|\widehat{S})$, where $f(x|S)$ is the conditional distribution of data given the unordered support $S$,
$$
f(x|S) 
= \sum_{P\in\sigma(S)} f(x|P) \pi^{\text{ord}}(P|S) 
= \frac{1}{s!} \left(\sum_{P\in\sigma{(S)}} \prod_{i=1}^s {f_{i}(x(P(i)))}\right) \prod_{k\not\in S}{f_0(x(k))},
$$
where $\sigma(S)$ is the set of all permutations of the indices in the support set $S$.

Suppose that $\widehat{S}$ is {\em not} an oracle thresholding estimator, then there must be indices 
$j \in \widehat{S}$ and $j' \not \in \widehat{S}$ such that $x(j) < x(j')$.
We exchange the labels of $x(j)$ and $x(j')$, and form a new estimate 
$\widehat{S}\,' = \big(\widehat{S}\setminus\{j\}\big)\cup\{j'\}$.
Comparing the likelihoods under $\widehat{S}$ and $\widehat{S}\,'$, we have
\begin{align}
    f(x|\widehat{S}) - f(x|\widehat{S}\,') 
    &= \frac{1}{s!} \sum_{P\in\sigma{(\widehat{S})}} \prod_{i=1}^s {f_{i}(x(P(i)))} f_0(x(j'))\prod_{k\not\in \widehat{S}\cup\{j'\}}{f_0(x(k))} - \nonumber \\
    &\quad\quad\quad - \frac{1}{s!} \sum_{P'\in\sigma{(\widehat{S}')}} \prod_{i=1}^s {f_{i}(x(P'(i)))} f_0(x(j)) \prod_{k\not\in \widehat{S}'\cup\{j\}}{f_0(x(k))} \nonumber \\
    &= \frac{1}{s!} \left(\sum_{i=1}^s a_i  \Big(f_i(x(j)) f_0(x(j')) - f_i(x(j')) f_0(x(j))\Big) \right) \prod_{k\not\in \widehat{S}\cup\{j'\}}{f_0(x(k))}, \label{eq:MLR-optimality-proof}
    % = \log{f_\delta(x(j))} + \log{f_0(x(j'))} - \log{f_\delta(x(j'))} - \log{f_0(x(j))} 
\end{align}
where the last equality follows by first summing over all permutations fixing $P(i) = j$ and $P'(i) = j'$, and setting $a_i = \sum_{P\in\sigma{(\widehat{S}\setminus\{j\})}} \prod_{i'\neq i} {f_{i'} (x(P(i')))}$. Notice that the $a_i$'s are non-negative.

Since $x(j) < x(j')$, and since each of $\{f_0, f_{i}\}$ is an MLR family, we have
$$
\frac{f_i(x(j))}{f_0(x(j))} - \frac{f_i(x(j'))}{f_0(x(j'))} \le 0 \implies f_i(x(j)) f_0(x(j')) - f_i(x(j')) f_0(x(j)) \le 0.
$$
Using Relation \eqref{eq:MLR-optimality-proof}, we conclude that $f(x|\widehat{S}) \le f(x|\widehat{S}\,')$.
Continuing this way, we can successively improve the likelihood of every estimator until we arrive at an oracle thresholding estimator, proving the desired optimality.  Note that with the same argument, we obtain that any two oracle 
thresholding estimators have the same likelihood.
\end{proof}

We emphasize that under the MLR conditions in Proposition \ref{prop:optimal-oracle-procedures},
 the oracle thresholding procedures are in fact \emph{finite-sample optimal} in the 
above Bayesian context. Further, our setup allows for different alternative distributions, and relaxes the assumptions 
of \citet{butucea2018variable} when studying distributional generalizations, where the alternatives are assumed to be identically distributed.

It remains to understand when the key MLR property holds.  We elaborate on this question next.
Returning to the more concrete signal-plus-noise model \eqref{eq:model-additive}, it turns out that the error tail behavior is what determines the optimality of data thresholding procedures.
In this setting, log-concavity of the error densities is \emph{equivalent} to the MLR property (Lemma \ref{lemma:MLR-log-concavity}). This, in turn, yields the finite-sample optimality of data thresholding procedures (Proposition \ref{prop:log-concave}).

\begin{lemma} \label{lemma:MLR-log-concavity}
Let $\delta$ be the magnitude of the non-zero signals in the signal-plus-noise model \eqref{eq:model-additive} with positive error density $f_0$, and let $f_\delta(x) = f_0(x-\delta)$.
The family $\{f_\delta, \delta \in \R\}$ has the MLR property if and only if the error density $f_0$ is log-concave.
\end{lemma}

\begin{proof}%[Proof of Lemma \ref{lemma:MLR-log-concavity}]
Suppose MLR holds, we will show that $f_0(t) = \exp\{\phi(t)\}$ for some concave function $\phi$.
By the assumption of MLR, for any $x_1 < x_2$, setting $\delta_0 = 0$, and $\delta_1 = (x_2 - x_1)/2 > 0$, we have
\begin{equation*}
    \log{\frac{f_{\delta_1}(x_2)}{f_{\delta_0}(x_2)}}
    = \phi\left(\frac{(x_1+x_2)}{2}\right)- \phi(x_2) 
    \ge \phi(x_1)- \phi\left(\frac{(x_1+x_2)}{2}\right) 
    = \log{\frac{f_{\delta_1}(x_1)}{f_{\delta_0}(x_1)}}.
\end{equation*}
This implies that the log-density $\phi(t)$ is midpoint-concave, i.e., for all $x_1$ and $x_2$, we have,
\begin{equation}
    \phi\left(\frac{(x_1+x_2)}{2}\right) 
    \ge \frac{1}{2} \phi(x_1) + \frac{1}{2} \phi(x_2).
\end{equation}
For Lebesgue measurable functions, midpoint concavity is equivalent to concavity by the Sierpinki Theorem \citep[see, e.g., Sec I.3 of][]{donoghue2014distributions}. This proves the `only-if' part.

For the `if' part, when  $\phi(t) = \log{(f_0(t))}$ is log-concave, then for any $\delta_0 < \delta_1$, and any $x<y$, we have
\begin{equation} \label{eq:concavity-implies-MLR}
    \log{\frac{f_{\delta_1}(y)}{f_{\delta_0}(y)}} - \log{\frac{f_{\delta_1}(x)}{f_{\delta_0}(x)}}
    = \phi(y-\delta_1) - \phi(y-\delta_0) - \phi(x-\delta_1) + \phi(x-\delta_0) \ge 0,
\end{equation}
where the last inequality is a simple consequence of concavity (see Lemma \ref{lemma:four-point-concavity} below). This proves the `if' part.
\end{proof}
\begin{lemma} \label{lemma:four-point-concavity}
Let $\phi$ be any concave function on $\R$. For any $x<y\in\R$, and $\delta>0$ we have
$$
\phi(x) + \phi(y+\delta) \le \phi(y) + \phi(x+\delta).
$$
\end{lemma}

\begin{proof}%[Proof of Lemma \ref{lemma:four-point-concavity}]
Pick $\lambda = \delta/(y-x+\delta)$, by concavity of $f$ we have
\begin{equation} \label{eq:four-point-concavity-1}
    \lambda \phi(x) + (1-\lambda) \phi(y+\delta) 
    \le \phi(\lambda x + (1-\lambda)(y+\delta)) 
    = \phi(y),
\end{equation}
and
\begin{equation} \label{eq:four-point-concavity-2}
    (1-\lambda) \phi(x) + \lambda \phi(y+\delta)
    \le \phi((1-\lambda) x + \lambda(y+\delta)) 
    = \phi(x+\delta).
\end{equation}
Summing up \eqref{eq:four-point-concavity-1} and \eqref{eq:four-point-concavity-2} and we arrive at the conclusion as desired.
\end{proof}

Proposition \ref{prop:optimal-oracle-procedures} and Lemma \ref{lemma:MLR-log-concavity} yield immediately the following.

\begin{proposition} \label{prop:log-concave}
Consider the additive error model \eqref{eq:model-additive}, where the $\epsilon(i)$'s are
independent with common distribution $F$.  Let the signal $\mu$ have $s$ positive entries with magnitudes $0<\delta_1\le\ldots\le\delta_s$, located on $\{1,\ldots,p\}$ as prescribed in \eqref{eq:uniform-ordered}.

If $F$ has a positive, log-concave density $f$, then the support estimator % oracle thresholding procedure 
$$
 \widehat{S}^* := \{i\, :\, x(i)\ge x_{[s]}\}
$$ 
is  %a.s.\ unique and
finite-sample optimal in terms of Bayes risk in the sense of \eqref{eq:finite-sample-Bayes-optimal}.
\end{proposition} 
\begin{proof} The independence and the fact that the observations have densities implies the absence of 
ties among the order statistics $\{x_{[i]}\}$, with probability one.  Thus, the oracle thresholding procedure is a.s.\ unique 
and given by $\widehat{S}^* = \{i\, :\, x(i)\ge x_{[s]}\}$.  The result then follows from 
Proposition \ref{prop:optimal-oracle-procedures} and Lemma \ref{lemma:MLR-log-concavity}.
\end{proof}

\begin{remark}
Propositions \ref{prop:optimal-oracle-procedures} and \ref{prop:log-concave} show that under MLR (or equivalently, 
log-concavity of the errors in additive models), the oracle thresholding procedures are finite-sample optimal even in the case where the signals have different (positive) sizes. This fascinating property perhaps 
explains the success of the thresholding estimators.
\end{remark}

The assumption of log-concavity of the densities is compatible with the AGG model when $\nu\ge1$, as 
demonstrated in the next example.

\begin{example} \label{exmp:AGG-logconcave}
The generalized Gaussian density $f(x)\propto \exp\{-|x|^\nu/\nu\}$ is log-concave for all $\nu\ge1$.
Therefore in the additive error model \eqref{eq:model-additive}, according to Proposition \ref{prop:log-concave}, the oracle thresholding procedure is Bayes optimal in the sense of \eqref{eq:finite-sample-Bayes-optimal}.
\end{example}

\section{Bayes optimality of likelihood ratio thresholding}
\label{subsec:optimal-procedure-super-exponential}

When the MLR condition in Proposition \ref{prop:optimal-oracle-procedures} is violated, the oracle thresholding 
procedures can in fact be sub-optimal (see Example \ref{exmp:suboptimal-data-thresholding} and 
Section \ref{sec:sup-optimality-of-thresholding}, below).  
%

In this section, we demonstrate that thresholding the \emph{likelihood ratio} rather than signal values
yields the finite-sample Bayes optimal procedures.  We consider a special but sufficiently general case of signal 
models with equal densities.

Namely, let the observations $x(i)$, $i=1,\ldots,p$ have $s$ signals as prescribed in \eqref{eq:uniform-ordered} with {\em common} ``signal'' density $f_a$, and let the remaining $(p-s)$ locations have common ``error'' density $f_0$.
Define the likelihood ratios 
$$
L(i) := {f_a(x(i))}\big/{f_0(x(i))},
$$
and let $L_{[1]} \ge L_{[2]} \ge \ldots \ge L_{[p]}$ be the order statistics of the $L(i)$'s.

\begin{definition}[Oracle likelihood ratio thresholding] Recall that $\langle a_1,\cdots,a_s\rangle$ denotes
the vector of $a_i$'s arranged in a non-increasing order.  Any estimator $\hat S = \{i_1,\cdots,i_s\}$ such that
$$
\langle L(i_1),\cdots,L(i_s) \rangle = (L_{[1]},\cdots,L_{[s]}),
$$
will be referred to as an {\em oracle likelihood thresholding} estimator of the support $S$.
\end{definition}
\begin{proposition} \label{prop:likelihood-ratio-thresholding} Any oracle likelihood ratio thresholding 
procedure $\widehat{S}_{\text{LRT}}$ is finite-sample optimal in terms of Bayes risk. That is,
\begin{equation}
    \widehat{S}_{\text{LRT}} \in \argmin_{\widehat{S} \in {\mathcal S}} \E_\pi \P[\widehat{S} \neq S].
\end{equation}
for all $s$ and $p$, where the infimum on $\widehat{S}$ is taken over all support estimators of size $s$.
\end{proposition}

\begin{proof}%[Proof of Proposition \ref{prop:likelihood-ratio-thresholding}]
The proof is analogous to that of Proposition \ref{prop:optimal-oracle-procedures}.
We need to show that
$
\widehat{S}_{\rm LRT} \in \argmax_{S \in \mathcal{S}} f(x|S) \pi(S).
$
Since the distribution $\pi$ of the support $S$ 
is uniform (recall \eqref{eq:uniform}), it is equivalent to prove that 
$$
f(x|\widehat{S}_{\text{LRT}}) = \max_{S \in \mathcal {S}} f(x|S),
$$ 
where  $f(x|S)$ is the conditional distribution of the data given the unordered support $S$,
\begin{equation} \label{eq:likelihood-ratio-thresholding-proof}
    f(x|S) = \sum_P f(x|P) \pi^{\text{ord}}(P|S) = \prod_{j\in S} f_a(x(j)) \prod_{j\not\in S}{f_0(x(j))}.
\end{equation}
Suppose $\widehat S\in {\mathcal S}$ is {\em not} an oracle likelihood thresholding estimator. 
Then from the definition of the likelihood ratio thresholding procedure, there 
must be indices 
$j \in \widehat S$ and $j' \not \in \widehat{S}$ such that $L(j) < L(j')$.
If we exchange the labels of $L(j)$ and $L(j')$, that is, we form a new estimate 
$\widehat{S}\,' = \big(\widehat{S}\setminus\{j\}\big)\cup\{j'\}$,
comparing the log-likelihoods under $\widehat{S}$ and $\widehat{S}\,'$, we have
\begin{equation*}
    \log{f(x|\widehat{S})} - \log{f(x|\widehat{S}\,')} 
    = \log{f_a(x(j))} + \log{f_0(x(j'))} - \log{f_a(x(j'))} - \log{f_0(x(j))}.
\end{equation*}
By the definition of $L(j)$'s, and the order relations, we obtain
\begin{equation*}
    \log{f(x|\widehat{S})} - \log{f(x|\widehat{S}\,')} 
    = \log{L(j)} - \log{L(j')} > 0
\end{equation*}
This shows that $\widehat S$ cannot be Bayes optimal unless it is a likelihood thresholding estimator.
Note that with the same argument for every two likelihood thresholding estimators $\widehat S'$
and $\widehat S''$, we have $ f(x|\widehat{S}') = f(x|\widehat{S}'')$, proving the desired optimality.
\end{proof}

The characterization of optimal likelihood ratio thresholding procedures in Proposition \ref{prop:likelihood-ratio-thresholding} may not always yield practical estimators, as the density of the alternatives, and the number of signals $s$ 
are typically unknown. Still, some insights can be gained by virtue of Proposition \ref{prop:likelihood-ratio-thresholding}.
In particular, when MLR fails (or equivalently, when the errors in model \eqref{eq:model-additive} do not have log-concave densities), data thresholding is sub-optimal.

\begin{example}[Sub-optimality of data thresholding] \label{exmp:suboptimal-data-thresholding}
Let the errors have iid generalized Gaussian density with $\nu=1/2$, i.e., $\log{f_0(x)}\propto -x^{1/2}$. 
Let dimension $p=2$, sparsity $s=1$ with uniform prior, and signal size $\delta=1$.
That is, $\mathbb P[\mu = (0,1)^\mathrm{T}] = \mathbb P[\mu = (1,0)^\mathrm{T}] = 1/2$.
If the observations take on values $x = (x_1, x_2)^\mathrm{T} = (1,2)^\mathrm{T}$, we see from a comparison of the likelihoods (and hence, the posteriors),
$$
\log \frac{f(x|\{1\})}{f(x|\{2\})} = 2x_1^{1/2} + 2(x_2 - 1)^{1/2} - 2x_2^{1/2} - 2(x_1 - 1)^{1/2} = 4 - 2\sqrt{2} > 0,
$$
that even though $x_1<x_2$, the set $\{1\}$ is a better estimate of support than $\{2\}$, i.e., $\mathbb P[S=\{1\}\,\big|\,x] > \mathbb P[S=\{2\}\,\big|\,x]$.
\end{example}

This simple example shows that, in the case when the errors have super-exponential tails, the optimal procedures are in general \emph{not} data thresholding.
A slightly more general conclusion can be found in Corollary \ref{cor:log-convex}.

\section{Sub-optimality of data thresholding procedures}\label{sec:sup-optimality-of-thresholding}

We provide a slightly more general result on the sub-optimality of data thresholding procedures.

\begin{corollary} \label{cor:log-convex}
Consider the additive error model \eqref{eq:model-additive}.
Let the errors $\epsilon$ be independent with common distribution $F$.
Let each of the $s$ signals be located on $\{1,\ldots,p\}$ uniformly at random with equal magnitude $0<\delta<\infty$.
% , independently with probability $p_1,\ldots,p_s$.
% , {\color{red} so that we have $f_a = p_1f_1 + p_2f_2 + \ldots + p_sf_s$.}
Assume the errors $\epsilon(i)$'s are iid with density $f$ that is log-convex on $[K, +\infty)$, for some $K>0$.

If $\widehat{S}_{\rm opt}$ is the Bayes optimal (i.e., the oracle likelihood thresholding estimator), then, whenever 
$j\in\widehat{S}_{\text{opt}}$ for some $x(j) > K+\delta$, we must necessarily have $j'\in\widehat{S}_{\text{opt}}$ for 
all $j'$ such that $K+\delta \le x(j') < x(j)$.
\end{corollary}

Specifically, if there are $m$ observations exceeding $K+\delta$, with $m>s$, then the top $m-s$ observations will \emph{not} be included in the optimal estimator $\widehat{S}_{\text{opt}}$. This shows that, in the case when the 
errors have super-exponential tails, the optimal procedures are in general \emph{not} data thresholding.

\begin{proof}[Proof of Corollary \ref{cor:log-convex}]
Since the density of the alternatives $f_{a}(t) = {f(t-\delta)}$ is log-convex on $[K+\delta, \infty)$, by Relation \eqref{eq:concavity-implies-MLR} in the proof of Lemma \ref{lemma:MLR-log-concavity} and appealing to log-convexity (rather than log-concavity), the likelihood ratio
$$
L(j) := \frac{f_a(x(j))}{f_0(x(j))} %= \sum_{i=1}^s \frac{p_if_{i}(x(j))}{f_0(x(j))},
$$
is decreasing in $x(j)$ on $[K+\delta, \infty)$.  The claim follows from Proposition \ref{prop:likelihood-ratio-thresholding}.
\end{proof}

\begin{remark}[{\bf An open problem\!}\,] As we have seen, the thresholding estimators are no longer optimal 
in the additive model with error-densities heavier than exponential.   Thanks to Proposition 
\ref{prop:likelihood-ratio-thresholding}, the oracle likelihood thresholding procedures are promising alternatives
that can lead us to practical support estimators.

In the case where the signals have different sizes, however, the argument in the proof of Proposition \ref{prop:likelihood-ratio-thresholding} breaks downs since the signal densities need to be identical for Relation \eqref{eq:likelihood-ratio-thresholding-proof} to hold.  In such cases, the characterization of the optimal procedure is an open problem.
\end{remark}

% \begin{remark}
% In view of Theorem \ref{thm:necessary-strengthened}, Bonferroni's thresholding procedure $\widehat{S}_p^{\text{Bonf}}$ described in Theorem \ref{thm:sufficient} is an asymptotically minimax procedure over the class of error models $U(F)$.
% That is, in the class of errors with AGG($\nu$) marginals $F$ where $\nu\ge1$, we have
% \begin{equation} \label{eq:minimax}
%     \widehat{S}_p^{\text{Bonf}} \in \argmin_{\widehat{S}_p} \sup_{{\cal E}\in U(F)} R(\widehat{S}_p),
% \end{equation}
% where the minimum is taken over \emph{all} procedures $\widehat{S}$.
% It can be shown that the supremum in the Bayes risk formulation \eqref{eq:minimax}, where $S$ is random and uniformly distributed, can be replaced by a supremum over all possible (deterministic) configurations of support sets.
% This can be done by applying Theorem 1.1 in the supplement of \citep{butucea2018variable}.
% We content ourselves with a Bayesian minimax result here.
% \end{remark}

\section{Minimax optimality in exact support recovery} \label{sec:minimax}

We establish in this section minimax versions of our results from Chapter \ref{chap:exact-support-recovery}.
Specifically, if we restrict ourselves to \emph{the class of thresholding procedures} $\cal T$ (defined in \eqref{eq:thresholding-procedure}), then Bonferroni's procedure is minimax optimal, for \emph{any} fixed dependence structure in the URS class.
This is formalized in Corollary \ref{cor:point-wise-minimax} below.% in Section \ref{subsec:point-wise-minimax}.
We refer to this result as \emph{point-wise} minimax, to emphasize the fact that this optimality holds for every \emph{fixed} URS array.

Meanwhile, if we search over \emph{all procedures}, but expand the model space to include {\em all} dependence 
structures, then a different minimax optimality statement holds for Bonferroni's procedure.
This result, formally stated in Section \ref{subsec:minimax-over-dependence}, is a consequence of our characterization 
of the finite-sample Bayes optimality of thresholding procedures in Section \ref{subsec:optimal-procedure-sub-exponential}.
% The result is of independent interest and it is the main contribution of this section.

\subsection{Point-wise minimax optimality for thresholding procedures}
\label{subsec:point-wise-minimax}

Theorems \ref{thm:sufficient} and \ref{thm:necessary} can be cast in the form of an asymptotic minimax statement. 
% (concealing the gap between sufficient and necessary conditions as pointed out in Remark \ref{rmk:gap-between-sufficient-necessary}).

\begin{corollary}[Point-wise minimax]
\label{cor:point-wise-minimax}
Let $\widehat{S}^{\text{Bonf}}$ be the sequence of Bonferroni's procedure described in Theorem \ref{thm:sufficient}. 
Let also the errors have common $\text{AGG}(\nu)$ distribution $F$ with parameter $\nu>0$, and $\Theta_p^+$ be as defined in \eqref{eq:minimax-signal-config-over}.
If $\underline{r}>g(\beta)$, then we have
\begin{equation} \label{eq:point-wise-minimax-above}
    \limsup_{p\to\infty} \sup_{\mu\in\Theta_p^+(\beta, \underline{r})} \P(\widehat{S}^{\text{Bonf}}_p \neq S_p) = 0,
\end{equation}
for arbitrary dependence structure of the error array ${\cal E} = \{\epsilon_p(i)\}_p$.
Let $\cal T$ be the class of thresholding procedures \eqref{eq:thresholding-procedure}. 
If $\underline{r}<g(\beta)$, then we have
\begin{equation} \label{eq:point-wise-minimax-below}
    \liminf_{p\to\infty} \inf_{\widehat{S}_p\in \cal T} \sup_{\mu\in\Theta_p^+(\beta, \underline{r})} \P(\widehat{S}_p \neq S_p) = 1,
\end{equation}
for any error dependence structure such that ${\cal E}\in U(F)$ and ${(-\cal E)}\in U(F)$.
\end{corollary}

\begin{proof}%[Proof of Corollary \ref{cor:point-wise-minimax}]
The first conclusion \eqref{eq:point-wise-minimax-above} is a restatement of Theorem \ref{thm:sufficient}.

For the second statement \eqref{eq:point-wise-minimax-below}, since $\underline{r}<g(\beta)$, we can pick a sequence $\mu^*\in\Theta_p^+(\beta, \underline{r})$ such that $|S_p| = \lfloor p^{1-\beta}\rfloor$, with signals having the same signal size $\mu(i)=(2r\log{p})^{1/\nu}$ for all $i\in S_p$, where $\underline{r}<{r}<g(\beta)$.
For this particular choice of $\mu^*$ we have $\mu^*\in\Theta_p^-(\beta, \overline{r})$ where $r<\overline{r}<g(\beta)$,
and according to Theorem \ref{thm:necessary}, we obtain $\lim_{p\to\infty} \inf_{\widehat{S}_p\in \cal T} \mathbb P[\widehat{S}_p \neq S_p] = 1$, for all dependence structures in the URS class.
\end{proof}

\begin{remark}
Theorem \ref{thm:necessary} is a stronger result than the traditional minimax claim in Relation \eqref{eq:point-wise-minimax-below}.
Indeed,  \eqref{eq:classification-impossible-dependent} involves an infimum (over the class $\Theta^-_p$) while \eqref{eq:point-wise-minimax-below} has a supremum (over the class $\Theta^+_p$).

On the other hand, Corollary \ref{cor:point-wise-minimax} is more informative than many minimax-type statements, since it applies ``point-wise'' to any fixed error dependence structure in the URS class.
\end{remark}

\begin{remark}
Corollary \ref{cor:point-wise-minimax} echoes Remark \ref{rmk:dependence-assumptions}: for a very large class of dependence structures, we cannot improve upon Bonferroni's procedure in exact support recovery problems (asymptotically), unless we look beyond thresholding procedures.
% There is a limited amount of literature on non-thresholding procedures that utilize the error dependence structures to improve power.
% A notable effort was made by \citet{jin2014optimality}, where the performance (in terms of Hamming loss) of a so-called graphlet screening procedure was studied. 
\end{remark}

% This is an enhancement (in the asymptotic sense) of the minimax results in Section 4.1 of \cite{butucea2018variable}, where the supremum was taken over the dependence structures.

\subsection{Minimax optimality over all procedures}
\label{subsec:minimax-over-dependence}

%(This also explains the $\gamma\ge 1$ condition in Theorem 1 of Arias-Castro and Chen)

Consider the asymptotic Bayes risk as defined in \eqref{eq:Bayes-risk}. The statement for the 
necessary condition of support recovery in Theorem \ref{thm:necessary}, with the help of Proposition 
\ref{prop:log-concave}, can be strengthened to include all procedures (in the Bayesian context), regardless 
of whether they are thresholding or not.

\begin{theorem} \label{thm:necessary-strengthened}
Consider the additive model \eqref{eq:model-additive} where the $\epsilon_p(i)$'s are independent and identically distributed with log-concave densities in the AGG class. 
Let the signals be as prescribed in Proposition \ref{prop:log-concave}.
If the signal sizes fall below the strong classification boundary \eqref{eq:strong-classification-boundary}, 
i.e. $\overline{r}<g(\beta)$, then we have
\begin{equation} \label{eq:classification-impossible-Bayes}
    \liminf_{p\to\infty} \inf_{\widehat{S}_p} \E_\pi \P[\widehat{S}_p\neq S_p] = 1,
\end{equation}
where the infimum on $\widehat{S}_p$ is taken over all procedures.
\end{theorem}

\begin{proof}%[Proof of Theorem \ref{thm:necessary-strengthened}]
When the errors are independent with log-concave density, the oracle thresholding procedure $\widehat{S}^*_p$, by Proposition \ref{prop:log-concave}, minimizes the Bayes risk \eqref{eq:Bayes-risk} among \emph{all} procedures. That is,
$$
\liminf_{p\to\infty} \inf_{\widehat{S}_p} \E_\pi\P[\widehat{S}_p\neq S_p]
\ge \liminf_{p\to\infty} \E_\pi\P[\widehat{S}^*_p\neq S_p].
$$
Since $\widehat{S}^*_p$ belongs to the class of all thresholding procedures, we have
\begin{align*}
    \liminf_{p\to\infty} \E_\pi\P[\widehat{S}^*_p\neq S_p] 
    &\ge \liminf_{p\to\infty} \inf_{\widehat{S}_p\in{\cal T}} \E_\pi\P[\widehat{S}_p\neq S_p] \\
    &\ge \liminf_{p\to\infty} \inf_{\widehat{S}_p\in{\cal T}} \inf_{S_p} \P[\widehat{S}_p\neq S_p] = 1,
\end{align*}
when $\overline{r}<g(\beta)$, where the last line follows from Theorem \ref{thm:necessary}.
\end{proof}

% In the previous section we demonstrated the minimax optimality of thresholding procedures, in a point-wise sense over a large class of dependence structure.
Theorem \ref{thm:necessary-strengthened} allows us to state another minimax conclusion --- one in which we search over \emph{all procedures}, by allowing the supremum in the minimax statement to be taken over the dependence structures.

\begin{corollary}
\label{cor:minimax-over-dependence}
Let $D(F)$ be the collection of error arrays with common marginal $F$ as defined in \eqref{eq:common-marginal-distribution} where $F$ is an $\text{AGG}(\nu)$ distribution.
Let also $\widehat{S}^{\text{Bonf}}_p$ be Bonferroni's procedure as described in Theorem \ref{thm:sufficient}.
If $\underline{r}>g(\beta)$, then we have
\begin{equation} \label{eq:minimax-over-dependence-over}
    \limsup_{p\to\infty} \sup_{\substack{\mu\in\Theta_p^+(\beta,\underline{r})\\ {\cal E}\in D(F)}} \P(\widehat{S}^{\text{Bonf}}_p \neq S_p) = 0.
\end{equation}
Further, when $\underline{r}<g(\beta)$, and $F$ has a positive log-concave density $f$, we have
\begin{equation} \label{eq:minimax-over-dependence-under}
    \liminf_{p\to\infty} \inf_{\widehat{S}_p} \sup_{\substack{\mu\in\Theta_p^+(\beta,\underline{r})\\ {\cal E}\in D(F)}} \P(\widehat{S}_p \neq S_p) = 
    1,
\end{equation}
where the infimum on $\widehat{S}_p$ is taken over all procedures.
\end{corollary}
\begin{proof}%[Proof of Corollary \ref{cor:minimax-over-dependence}]
Relation \eqref{eq:minimax-over-dependence-over} is a re-statement of Remark \ref{rmk:sufficient-strengthened}.

For any distribution $\pi$ (with a slight abuse of notation) over the parameter space $\Theta_p^+\times D(F)$, we have
\begin{equation} \label{eq:minimax-over-dependence-under-proof}
    \liminf_{p\to\infty} \inf_{\widehat{S}_p} \sup_{\substack{\mu\in\Theta_p^+(\beta, \underline{r})\\ {\cal E}\in D(F)}} \P(\widehat{S}_p \neq S_p) 
\ge \liminf_{p\to\infty} \inf_{\widehat{S}_p} \E_{\pi} \P(\widehat{S}_p \neq S_p),
\end{equation}
since the supremum is bounded from below by expectations.
In particular, define $\pi$ to be the uniform distribution over the configurations $\Theta_p^*\times I(f)$, where
\begin{align*}
    \Theta_p^* &=
    \{\mu\in\mathbb{R}^d:\;|S_p|=\lfloor p^{1-\beta}\rfloor,\;\mu(i)=0\;\text{for all }i\not\in S, \;\text{and}\\
    &\quad\quad\mu(i) = (\nu{r}\log{p})^{1/\nu}\;\text{for all }i\in S,\;\text{where}\;\underline{r}<r<g(\beta)\},
\end{align*}
and 
% \begin{align*}
%     I(f)&=\{{\cal E}=(\epsilon_p(i))_p:\;\epsilon_p(i)\;\text{are independently and identically distributed}\\
%     &\quad\quad\text{with density}\; f(x)\propto \exp\{-|x|^\nu/\nu\}\}.
% \end{align*}
\begin{equation*}
    I(f)=\{{\cal E}=(\epsilon_p(i))_p:\;\epsilon_p(i)\;\text{\ac{iid} with density}\; f(x)\propto \exp\{-|x|^\nu/\nu\}\}.
\end{equation*}

Since the density $f$ of $F$ is log-concave, the distribution of the signal configurations satisfies the conditions in Theorem \ref{thm:necessary-strengthened}.
Thus, the desired conclusion \eqref{eq:minimax-over-dependence-under} follows from Theorem \ref{thm:necessary-strengthened} and \eqref{eq:minimax-over-dependence-under-proof}.
\end{proof}

\begin{remark}
Since the class $\text{AGG}(\nu)$, $\nu\ge1$ contains distributions with log-concave densities (Example \ref{exmp:AGG-logconcave}), the minimax statement \eqref{eq:minimax-over-dependence-under} continues to hold if the supremum is taken over the entire class $F\in\text{AGG}(\nu)$, $\nu\ge1$.
We opted for a more informative formulation which emphasizes the log-concavity condition on the density of $F$.
\end{remark}

\begin{remark}
Corollary \ref{cor:minimax-over-dependence} is no stronger than Corollary \ref{cor:point-wise-minimax}. 
In Corollary \ref{cor:point-wise-minimax} we search over only the class of thresholding procedures, but offer a tight, point-wise lower bound on the asymptotic risk over the class of URS dependence structures.
On the other hand, Corollary \ref{cor:minimax-over-dependence} provides a uniform lower bound for the asymptotic risk over all dependence structures, which may not be tight except in the case of independent errors.
% , and when the errors have nice, log-concave densities.
\end{remark}

\section{Optimality and sub-optimality: A discussion}

We conclude with a brief summary on the optimality and sub-optimality of the thresholding procedures in the problem of exact support estimation.
For clarity, we focus on the model \eqref{eq:model-additive} with {\em independent} errors.

Theorem \ref{thm:necessary-strengthened} and Corollary \ref{cor:minimax-over-dependence} provide a nearly complete picture of the difficulty 
in the exact support  recovery problem, in the regime when the thresholding estimators are optimal. Specifically, in such cases the signal classification boundary is universal.  On the other hand, Proposition \ref{prop:likelihood-ratio-thresholding}, and indeed, Example \ref{exmp:suboptimal-data-thresholding} demonstrate that thresholding procedures are \emph{sub-optimal} for $\text{AGG}(\nu)$ models with $\nu<1$. Therefore, the optimality of thresholding
procedures (specifically, Bonferroni's procedure) only applies to $\text{AGG}(\nu)$ models with $\nu\ge1$. 
% This suggests that the condition $\nu\ge1$ in Corollary \ref{cor:minimax-over-dependence} is tight.

If we restrict the space of methods to only thresholding procedures, then the results in Section \ref{subsec:point-wise-minimax} state that the phase 
transition phenomenon --- the 0-1 law in the sense of Corollary \ref{cor:point-wise-minimax} --- is universal in all error models with rapidly varying tails.
This includes $\text{AGG}(\nu)$ models {\it for all} $\nu>0$.  In contrast, models with heavy (regularly varying) tailed errors do not exhibit this phenomenon 
(form more details, see Theorem \ref{thm:heavy-tails}).  We summarize the properties of thresholding procedures in 
Table \ref{table:role-of-thresholding}.

%\end{remark}

\begin{table}[ht]
    \centering
    \caption{Properties of thresholding procedures under different error distributions when the errors are independent. 
    Properties of the error distributions are listed in brackets.}
    \medskip
    \begin{tabular}{p{40mm}p{37mm}p{37mm}} \toprule
        Thresholding procedure & Bayes optimality &  Phase transition \\ 
        (Error distributions) &  (Log-concave density) & (Rapidly-varying tails) \\ \midrule
        AGG($\nu$), $\nu\ge1$ & Yes (Yes) & Yes (Yes) \\ \cmidrule{1-3}
        AGG($\nu$), $0<\nu<1$ & No (No) & Yes (Yes) \\ \cmidrule{1-3}
        Power laws & No (No) & No (No) \\ \bottomrule
    \end{tabular}
    \label{table:role-of-thresholding}
\end{table}

% It should be remarked that Corollary \ref{cor:log-convex} assumes the existence and knowledge of a maximum signal size $\delta_s$. 
% When maximum signal sizes are unknown a priori, a large observation may be attributed to either a truly large signal, or to the error.
% Common practice has been to attribute large observations to signals, and not errors.

% As an example, consider the linear regression
% \begin{equation} \label{eq:regression}
%  Y = X\mu + \xi,
% \end{equation}
% where $\mu$ is a vector of regression coefficients of interest to be inferred from observations of $X$ and $Y$.
% If the design matrix $X$ is of full column rank, then the OLS estimator of $\mu$ can be formed 
% \begin{equation*}
%     \widehat{\mu} = \left(X'X\right)^{-1}X'Y = \mu + \epsilon,
% \end{equation*}
% where $\epsilon := (X'X)^{-1}X'\xi$.
% Hence we recover the generic problem \eqref{eq:model}. 
% The support recovery problem is therefore equivalent to the fundamental model selection problem.
% Often an investigator calculates the t-scores of each coefficient as 
% \begin{equation} \label{eq:linear-model-selection}
%     \widehat{\mu}(i) \Big/ \widehat{\mathrm{s.e.}}(\widehat{\mu}(i)),
% \end{equation}
% where $\widehat{\mathrm{s.e.}}(\widehat{\mu}(i))$ is the estimated standard error of $\widehat{\mu}(i)$.
% The investigator then chooses indices with large t-scores to enter the model.
% If the errors in model \eqref{eq:regression} are iid Gaussian, the expression \eqref{eq:linear-model-selection} is t-distributed and have power-law tails; the discussion above suggests that this commonplace procedure may be sub-optimal for bounded signals.