### 2021-02-04 preamble.tex

\preface

%\fbox{From Springer}
%{\color{blue}
%% Please write your preface here
%Use the template \emph{preface.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to %style your preface in the Springer layout.

%A preface\index{preface} is a book's preliminary statement, usually written by the \textit{author or editor} of a work, which states its origin, scope,
 %purpose, plan, and intended audience, and which sometimes includes afterthoughts and acknowledgments of assistance. 
%When written by a person other than the author, it is called a foreword. The preface or foreword is distinct from the introduction, which deals with 
%the subject of the work.
%
%Customarily \textit{acknowledgments} are included as last part of the preface.}

This text presents a collection of new results and recent developments on the phase transition phenomena in sparse signal problems.
The main theme is the study of the fundamental limits in high-dimensional testing and inference. Since the seminal works
of \cite{ingster1998minimax} and \cite{donoho2004higher}, the subject has received a lot of attention in the literature with important contributions
from \cite{ji2012ups,genovese2012comparison,jin2014optimality,arias2017distribution,butucea2018variable}.  These works, among many others, have discovered some fundamental limits in the so-called  {\em needle in haystack} problems, where a sparse signal is observed with high-dimensional additive noise.  In this setting, two archetypal  problems arise -- the {\em signal detection} and {\em signal support recovery}.  
The signal detection refers to a global hypothesis testing problem that amounts
to determining the presence of non-zero signal in any of its dimensions. 
The support recovery, on the other hand, can be seen either as a multiple testing problem where the presence of non-zero signal is tested for each signal location of interest, or alternatively,  as an inference problem that aims to estimate the signal support, i.e., the locations of the non-zero signal components.  The fundamental limits of these problems are studied in the so-called high-dimensional asymptotic regime where the dimension $p$ of the underlying signal grows to infinity, and the sample size $n$ is either bounded or grows slowly relative to $p$.

From a probabilistic perspective, these aforementioned fundamental limits are stated as asymptotic zero-one type laws, as dimensionality diverges.  
Namely, consider a sparse signal with  {\em support size} on the order of $p^{1-\beta}$ for some parameter $\beta\in (0,1)$.  
Parameterize the non-zero {\em signal amplitude} by $\Delta(p^r)$, 
for some $r>0$ and a suitable monotone non-decreasing function $\Delta(\cdot)$.  Then, for a broad range of error distributions and
statistical problems, one encounters a sharp transition between the regimes where the problem is solvable and unsolvable depending on 
the signal magnitude $r$ and signal sparsity $\beta$.  More precisely, there exists a boundary function $f(\beta)$ such that if 
the signal magnitudes are {\em above} the boundary, $r> f(\beta)$, then the problem can be solved with vanishing loss as $p\to\infty$, with a suitable statistical procedure.  On the other hand, if the signal is below that same boundary, i.e., $r<f(\beta)$, all statistical procedures fail to  provide a solution with a vanishing loss, as $p\to\infty$. Of course, depending on whether one considers the  detection (testing) or 
support recovery (inference) problems, different loss functions quantify success and failure. The choice of the loss functions is often guided by the 
applications, resulting in a rich picture of phase-transitions (see e.g.\ Figure \ref{fig:phase-Gaussian-errors}).

{\bf The contributions of this work.} The fundamental limits of the classic detection problem hinge 
of the analysis of the discrepancy between the {\em null} and {\em alternative} hypotheses e.g., via Hellinger distance.  Thus, perhaps 
for  technical reasons, much of the analysis in the existing literature has been done under the assumption that the additive errors are 
independent and/or Gaussian, or using loss functions unaffected by the dependence such as the Hamming loss.  
In this work, we demonstrate that the support recovery problems, especially {\em exact support recovery}, are best understood from 
the novel perspective of the {\em concentration of maxima} phenomenon in extreme value theory.  It turns out that under a very broad 
range of light-tailed error distributions and under a {\em very} broad range of error dependence structures, the maxima of the errors, 
when rescaled (but not centered!) converge in probability to a positive constant. This concentration property leads to a complete solution 
of the exact  support recovery problem for the broad family of thresholding procedures.  Most if not all existing support estimation 
procedures are types thresholding procedures (see Section \ref{sec:statistical-procedures}).   
That is, the signal support estimate comprises of all components exceeding a suitable (potentially data-dependent)  threshold.  We show, by exploiting concentration of maxima, that thresholding procedures obey a phase-transition, where if the signal is above a 
certain boundary, asymptotically exact recovery is possible while below the boundary all thresholding procedures fail, as $p\to\infty$.  
Remarkably, light-tailed maxima concentrate under very broad and strong dependence.  This is exemplified by our characterization of
the concentration of maxima phenomenon for Gaussian triangular arrays.  For example, in the special case of stationary Gaussian time
series, vanishing auto-covariance is necessary and sufficient for the maxima to concentrate in the same way as independent standard 
normal random variables.  This is in stark contrast with the behavior of sums, commonly studied under short- and long-range dependence conditions \citep[see e.g.][]{dedecker:doukhan:etal:2007,pipiras:taqqu:2017}.  Simply put, the notion of weak dependence that entails that the maxima of dependent variables concentrate at the
same rate as in the case of independence is fundamentally weaker than the conventional mixing conditions widely used in the study of sums.

Our probabilistic contributions may be of independent interest and extend classic work of \cite{berman1964limit}. 
Concentration of maxima is a type of super-concentration phenomenon studied also in \cite{chatterjee2014superconcentration} 
and \cite{tanguy:2015}. The robustness of the concentration of maxima phenomenon to dependence can perhaps explain the universality 
of phase transitions in support recovery problems.

The use of concentration of maxima phenomenon highlights one core idea in our work, which allows for a first of its kind comprehensive 
treatment of thresholding procedures under very broad error-dependence conditions.  The text involves also a full spectrum of related results 
such as minimax-optimality and finite-sample Bayes optimality in support estimation.  Using different type of loss functions and type I error 
controls, we obtain a rich picture of the exact and approximate support recovery problems in high dimensions.  
Many of these phase transition results have not appeared in previously published literature.

High-dimensional support recovery problems arise in many modern applications ranging from cybersecurity, theoretical computer science,
to statistical genetics.  Genome-wide association studies (GWAS) in genetics are particularly natural applications, where the asymptotic 
phase-transition results help explain and quantify a previously observed empirical phenomenon of the so-called {\em steep part of the power 
curve}  In the last chapter of this work, we detail this application and highlight future theoretical and practical consequences of our work.

{\bf Target audience.} The original research presented in this text originates from the doctoral dissertation of the first author in the 
Statistics department at the University of Michigan, Ann Arbor.  The main goal of this text is to provide a comprehensive treatment of the
exact and approximate support recovery problems by utilizing existing and newly developed probabilistic tools on concentration of maxima.
The text also provides a quick introduction to the state-of-the-art in the dynamic area of phase-transitions in high-dimensional 
testing and inference.  It is accessible to doctoral students in Statistics with background in measure-theoretic probability and  
statistics as well as to researchers in applied fields working with high-dimensional data sets.  The text can be used as a reference and a
supplement to a special topics course on high-dimensional inference.

{\bf Acknowledgements.} The authors gratefully acknowledge the support of their families and all colleagues from the Statistics Department 
at the University of Michigan, Ann Arbor.   Special thanks (in alphabetical order) go to Xuming He, Tailen Hsing, Michalis Kallitsis, Liza Levina, Yuanzhi Li, Rodderick Little, Ya'acov Ritov, Kerby Shedden, Jinqi Shen, Jonathan Terhorst, Gongjun Xu.
The authors were partially supported by the NSF program {\em Algorithms for Threat Detection}.

\vspace{\baselineskip}
\begin{flushright}\noindent
Chicago \hfill {\it Zheng Gao }\\
Ann Arbor, \DateEdited \hfill {\it Stilian Stoev}\\
\end{flushright}