### 2021-03-10 7.GWAS/GWAS.tex

The process of scientific discovery, as famously explained by Richard Feynman, usually starts with guesses in one form or another.
The consequences of such guesses are then computed, and compared with experimental results.
If the predictions disagree with the experiment, then our guesses are wrong. ``That is all there is to it''  \citep{feynman2017character}.

In the previous chapters, we delved deep into the theoretical underpinnings of the phase transition phenomena in high-dimensional multiple testing problems.
The results are interesting in their own right.
However, in the spirit of Feynman, we have not discovered any scientific law, but merely worked out mathematical consequences of our postulated models.
In this chapter, our goal is to relate these predictions to real experimental data from the field of genetics, where large-scale simultaneous hypotheses testing problems often arise.
From such comparisons, we will demonstrate that the phase transition laws are indeed accurate predictions of some curious phenomena in the real world.
The accuracy of our predictions will lend credibility to the application of these ``laws of large dimensions'' in actual statistical problems.

\medskip

In our case, the experimental data used as the measuring stick come from genome-wide association studies (\ac{GWAS}), introduced in Section \ref{sec:motivation-chisq}.
Recall that in \ac{GWAS}, a large number of marginal association tests are conducted simultaneously, resulting in statistics that can be approximated by
\begin{equation} \label{eq:model-chisq-Chapter6}
    %x(i) \distras{\mathrm{ind.}} \chi_\nu^2\left(\lambda(i)\right), \quad i=1,\ldots,p.
    x(i) \sim \chi_\nu^2\left(\lambda(i)\right), \quad i=1,\ldots,p,
\end{equation}
where $\chi_\nu^2\left(\lambda(i)\right)$ is a chi-square distributed random variable with $\nu>0$ degrees of 
freedom\footnote{The parameter $\nu$ here should not be confused with the shape parameter of the AGG$(\nu)$ 
distributions, which will not appear in this chapter.} and non-centrality parameter $\lambda(i)$.

We establish our theoretical predictions in two steps.
In Section \ref{sec:chisq-boundaries} below, we shall first establish the phase transitions of the model \eqref{eq:model-chisq-Chapter6}.
In parallel to results in Chapter \ref{chap:phase-transitions}, we show that several commonly used family-wise error rate-control procedures --- including Bonferroni's procedure --- are asymptotically optimal for the {exact}, and {exact-approximate} support recovery problems (recall Definition \ref{def:exact-recovery-success-failure}) in idealized chi-square models with independent components.
Analogously, the \ac{BH} procedure is asymptotically optimal for the {approximate}, and {approximate-exact} support recovery problems.
% \ref{thm:chi-squared-exact-boundary}, \ref{thm:chi-squared-exact-approx-boundary}, \ref{thm:chi-squared-approx-boundary}, and \ref{thm:chi-squared-approx-exact-boundary}.
Under appropriate parametrization of the signal sizes and sparsity, we establish the phase transitions of support recovery problems in the chi-square model.
Remarkably, the degree-of-freedom parameter does not affect the asymptotic boundaries in any of the four support recovery problems.

We then translate the canonical signal size and sparsity parametrizations into the vernacular of statistical geneticists in Section \ref{sec:odds-and-power}.
% present empirical evidence for the phase transition in the exact-approximate problem using real data from large-scale association studies on breast cancer obtained from the NHGRI-EBI GWAS Catalog \citep{macarthur2016new}.
% demystify the notion of signal size $\lambda$ in this context.
% which is perhaps less transparent than in additive error models.
We do so by characterizing the relationship between the signal size $\lambda$ and the marginal frequencies, odds ratio, and sample sizes for association tests on 2-by-2 contingency tables. 
These parameters are more often estimated and reported in \ac{GWAS} than the elusive signal size parameter $\lambda$.
% Specifically, the amount of signal, when rare variants are present, is weaker compared to the signal when the marginal distributions are balanced.
% In other words, reliable detection of the effects by rare variants would require more samples compared to common variants, even at the same odds ratio.
% This result, establishing the relationship between sample sizes and signal sizes, is made precise in Section \ref{sec:odds-and-power}.
As a bonus, we point out the implications of this relationship on statistically optimal study designs for association studies in Section \ref{sec:optimal-design}.
Perhaps surprisingly, our analysis reveals that balanced designs with equal number of cases and controls are often statistically inefficient.
% Practical consequences of these results in power analysis will be illustrated with data examples in Section \ref{sec:phase-transitions-in-GWAS}.

Armed with the results on phase transitions in the chi-square model, and a translation from the language of high-dimenstional statistics to the {\em patois} of association screening studies, we finally present in Section \ref{sec:phase-transitions-in-GWAS} the consequences of the phase transitions in \ac{GWAS}, and compare against real experimental data to evaluate the success of our predictions.

The phase transitions in the chi-square models are demonstrated with numerical simulations in Section \ref{sec:numerical}.
The proofs, which are closely related to those of the results in Chapter \ref{chap:phase-transitions},
 are collected in Section \ref{sec:supplement:proofs}.
%in Section \ref{sec:proof-signal-size-odds-ratio}.

% Practical issues are also addressed to make for simple and effective power analysis.

\section{Support recovery problems in chi-squared models}
\label{sec:chisq-boundaries}
\input{7.GWAS/7.chi-square.tex}

\section{Odds ratios and statistical power}
\label{sec:odds-and-power}
\input{7.GWAS/7.odds-and-power.tex}

\section{Optimal study designs and rare variants}
\label{sec:optimal-design} 
\input{7.GWAS/7.optimal-design.tex}

\section{Phase transitions in large-scale association screening studies}
\label{sec:phase-transitions-in-GWAS}
\input{7.GWAS/7.GWAS-phase-transitions.tex}

\section{Numerical illustrations of the phase transitions in chi-square models}
\label{sec:numerical}
\input{7.GWAS/7.numerical.tex}

%\section{Proofs}
%\label{sec:proof-signal-size-odds-ratio}
%\input{6.GWAS/6.proofs-GWAS.tex}