### 2021-04-05 4.exact_support_recovery/exact_support_recovery.tex

We focus on exact support recovery problems in this chapter.
%Here, we limit ourselves to the general class of thresholding estimators but allow for nearly arbitrary error dependence structure.  
%Questions on optimality and extensions to other classes of estimators are pursued in Chapter \ref{chap:optimality}.  
Recall from Lemma \ref{lemma:risk-exact-recovery-probability} that in order to study the asymptotic behaviors of $\mathrm{risk}^{\mathrm{E}}$, it is sufficient to establish minimal conditions under which the support sets can be consistently estimated, i.e.,
\begin{equation} \label{eq:exact-recovery} 
\P[\widehat{S}_p = S_p] \longrightarrow 1 \quad \text{as } \; p\to\infty, 
\end{equation}
where $\widehat{S}_p$ is an estimate of the true support set $S_p$ of a high-dimensional signal vector $\mu_p$.

We will establish minimal conditions such that \eqref{eq:exact-recovery} holds, by generalizing the results we obtained in Chapter \ref{chap:phase-transitions} to additive error models with relaxed distributional and dependence assumptions on the additive error array.

%Sufficient conditions for exact support recovery are given in Section \ref{subsec:sufficient}.  A very general class of 
%dependence structures characterized by the uniform relative stability concept will be introduced in Section \ref{subsec:URS}, to prepare us 
%for the necessary condition in Section \ref{subsec:necessary}.
%Section \ref{subsec:dense-signals} discusses the dense signal regime, and Section \ref{suppsec:numerical} illustrates the phase transition 
%phenomena with numerical examples.  We begin with an overview of our results in the context of the existing literature.

\section{Generalizations of distributional and dependence assumptions}
Consider the additive error model \eqref{eq:model-additive} with the triangular array of errors,
\begin{equation} \label{eq:error-array}
    {\cal E} = \left\{ (\epsilon_p(i))_{i=1}^p,\ p=1,2,\dots\right\},
\end{equation}
where the $\epsilon_p(i)$'s have common cumulative distribution function $F(x)=\P[\epsilon_p(i)\le x]$.
In contrast to the assumptions in Chapter \ref{chap:phase-transitions}, we only require the errors to have common marginal distributions.
% We will study the role of dependence in support recovery problems in Section \ref{sec:URS}.

Although our method of analysis applies to all light-tailed error distributions with rapidly varying tails (see Definition \ref{def:rapid-variation}), to be concrete and better convey the main ideas, we will focus on the class of $\mathrm{AGG}(\nu)$ laws (see Definition \ref{def:AGG}).
Extensions of the results to other classes of error models are presented in Section \ref{suppsec:other-boundaries}.

This generalized distributional assumption on the errors call for a suitable generalization of the signal size parametrization in order to analyze the problem as we did in the previous chapter.
As before, we assume the signals in model \eqref{eq:model-additive} to be a sparse vector $\mu = \left(\mu(i)\right)_{i=1}^p$ where its sparsity, with a few exceptions which will be explicitly stated, is parametrized as
\begin{equation} \label{eq:sparsity-parametrized}
   s = |S_p| = \lfloor p^{1 -\beta} \rfloor, % \quad \text{with } 0 < \beta \le 1 \; \text{(fixed)},
\end{equation}
with $0 < \beta \le 1$ fixed.

We assume that the non-zero entries of $\mu$ are positive and take values in the interval $\left[\underline{\Delta},\overline{\Delta}\right)\subset (0,\infty)$.
That is, $0<\underline{\Delta}\le\mu(i)<\overline{\Delta}\le+\infty$, for all $i\in S_p$.
The lower and upper bound on the signal sizes $\underline{\Delta}$ and $\overline{\Delta}$ are parametrized as
\begin{equation} \label{eq:signal-size-parametrized}
    \underline{\Delta} = \underline{\Delta}(p) = (\nu \underline{r} \log{p})^{1/\nu} \quad \text{and} \quad
    \overline{\Delta} = \overline{\Delta}(p)  = (\nu \overline{r} \log{p})^{1/\nu},
\end{equation}
with parameters $0 < \underline{r} \le \overline{r} \le +\infty$.
%The signal sizes now depends on the shape of the assumed error distributions $\mathrm{AGG}(\nu)$ through the parameter $\nu$.

 \stilian{\begin{remark}Without much effort, one can relax the above parameterization of the support size in \eqref{eq:sparsity-parametrized} to $|S_p| \sim p^{1-\beta}$ or in fact, to the 
much more general case of $|S_p|\sim L(p) p^{1-\beta}$, for a slowly varying function $L$.  Going along this path would would naturally lead to
more sophisticated signal models, which merit independent study.  Here, we opt for simplicity of exposition and thus adopt 
\eqref{eq:sparsity-parametrized} and \eqref{eq:signal-size-parametrized}.\end{remark}}

\medskip

We now turn to the dependence conditions.
Several authors have studied the support recovery problem in terms of the Hamming loss and obtained minimax optimality results (see, e.g., \cite{ji2012ups, genovese2012comparison, jin2014optimality, butucea2018variable}).
%% We briefly review the recent work by \cite{butucea2018variable}, whose .
In the special case of Gaussian marginals, \cite{butucea2018variable} showed that the boundary \eqref{eq:strong-classification-boundary-Gaussian} exists in a minimax sense.
That is, when the errors are \emph{independent} Gaussians, the Hamming loss cannot be made to vanish if the signal sizes are sufficiently small by any procedure. 
Conversely, if signal size falls below, the Hamming loss can be made to vanish for some thresholding procedure.
%
However, as pointed out in Section \ref{sec:risks-relations}, vanishing Hamming loss is only sufficient, not necessary for support recovery \eqref{eq:exact-recovery}, and results on the former do not carry over directly to the study of the exact support recovery problem.
%%More importantly, despite their elegance, these Hamming loss-minimax studies often amount to the analysis of the elementary case of iid data, and 
%%unless is by design blind to non-trivial error-dependence structures.
%%This prevents us from fully exploring of the phase transition phenomena under other dependence conditions.
More importantly, since Hamming loss decomposes into expectations on individual terms that are not affected by dependence, Hamming loss-minimax studies do not reveal the difference in probability of support recovery between independent and dependent observations. 
This prevents one from fully exploring the phase transition phenomena under other dependence conditions.
%% This is arguably a consequence of the inherent limitation of the Hamming-loss approach. 
%% For practitioners, a minimax statement in the Gaussian case seems to offer little guidance.
%% Indeed, in applications, independence is an exception rather than the rule; Gaussianity of the errors may also be unnecessarily restrictive.
As a result, so far in the literature, the role of dependence in model \eqref{eq:model-additive} have remained largely unexplored. 
%This chapter offers an advances in this direction, and provides a close-to-complete solution of the exact support recovery problem. (See also, Chapter \ref{chap:URS}.)
%We briefly summarize our contributions next.

We take a different approach in this text. 
In particular, we study the exact support recovery problem \eqref{eq:exact-recovery} directly, and show that for thresholding procedures the phase transition phenomena exists {universally} in a large class of dependence structures, and not just in a minimax sense.

In a first step, we show that in the AGG model under \emph{arbitrary} dependence, under the scaling described in \eqref{eq:sparsity-parametrized} and \eqref{eq:signal-size-parametrized}, the function
\begin{equation} \label{eq:strong-classification-boundary}
    f_{\mathrm{E}}(\beta) = f_{\mathrm{E},\nu}(\beta) = (1 + (1 - \beta)^{1/\nu})^\nu, \quad \nu>0,
\end{equation}
demarcates the region of possibility for the exact support recovery problem.
That is, if the signal sizes are above the boundary (i.e., $\underline{r}> f_{\mathrm{E}}(\beta)$), then \ac{FWER}-controlling procedures with appropriately calibrated levels achieve exact support recovery (Theorem \ref{thm:sufficient} below).
We refer to \eqref{eq:strong-classification-boundary} as the {\em strong classification boundary}.

Conversely, we show that for a surprisingly large class of dependence structures characterized by the concept of \emph{uniform relative stability} (URS, see Definition \ref{def:URS} below), when the signal size is below the boundary  (i.e., $r<f_{\mathrm{E}}(\beta)$), no thresholding procedure can achieve the asymptotically perfect support recovery. In fact,
\begin{equation} \label{eq:exact-recovery-failure}
    \mathbb{P}\left[\widehat{S}_p=S_p\right]\longrightarrow 0,\quad \mbox{ as }p\to \infty,
\end{equation}
for all thresholding procedures (Theorem \ref{thm:necessary} below).
These two results show that the thresholding procedures obey a phase transition phenomenon in a strong, \emph{point-wise} sense over the class of URS dependence structures, and over the class of AGG$(\nu),\ \nu>0$ error distributions. 
% A transparent characterizations of the dependence conditions under which this result holds will be established in the Chapter \ref{chap:URS} later.

%The conclusions are fundamentally stronger and more informative than the minimax statements in the literature (see, e.g., \cite{butucea2018variable}).
%The techniques developed in this here are also entirely different from those in \citet{ji2012ups} or \citet{butucea2018variable}, and transparent characterizations of the dependence conditions under which the phase transition type result holds will be established in the Chapter \ref{chap:URS} later.

%In Chapter \ref{sec:optimality} we study the finite-sample optimality of the threshold-based support estimation procedures in a Bayesian setting.

% The strong classification boundary $g$ characterizes a phase-transition phenomenon similar to that of the signal detection and approximate support recovery. 
% A preview of this result is presented in Figure \ref{fig:phase}.

% \begin{figure}
%       \centering
%       \includegraphics[width=0.45\textwidth]{./figures/phase_diagram_Gaussian_no_dashed_area.eps}
%       \includegraphics[width=0.45\textwidth]{./figures/phase_diagram_double_exponential.eps}
%       \caption{The phase diagrams of the detection, weak classification, and strong classification boundaries against sparse alternatives under Gaussian (left) and Laplace distributed (right) errors. Here $\beta$ and ${r}$ parametrize the signal sparsity and the lower and upper bounds of the signal sizes, respectively. 
%       The signal detection problem can be answered perfectly (asymptotically) inside the \emph{Detectable} region; false discovery proportion and non-discovery proportion can be made to vanish in the \emph{Weakly classifiable} region. We study in this paper the strong classification boundary, above which the support recovery can be achieved \emph{exactly} in the \emph{Classifiable} region $\{(\beta, {r}):{r}>g(\beta)\}$. 
%       In a large class of dependence structures characterized by URS, when signal sizes fall below the strong classification boundary \eqref{eq:strong-classification-boundary}, i.e. $\{(\beta, {r}):{r}<g(\beta)\}$, no thresholding procedure succeeds in the exact support recovery problem. }
%       \label{fig:phase}
% \end{figure}

%\medskip
%Our method of analysis is sufficiently general and can be applied to other error models. 
%The phase transition phenomena for two additional classes of error distributions with either heavier or lighter tails than the AGG distributions will be described in Section \ref{suppsec:other-boundaries}.

% In this sense, for $\nu\ge 1$ our results are rather complete.  We note here that our focus is not on minimax analysis of the support recovery problem, where a large class of distributional and dependence structures are considered.  
% In some sense, the worst case scenario (see also \cite{butucea2018variable}) is when the errors are independent.  
% In contrast, we establish the phase-transition phenomenon, for fixed but very general dependence conditions characterized by the URS property.

%We show that when the error-distribution $F$ is log-concave, then the thresholding procedures are optimal (for independent errors).  Therefore, in this regime the {\em strong classification boundary} $g$ is universal.  
%That is, if $r<g(\beta)$, no estimator can achieve perfect support recovery as $p\to\infty$. 
% \fbox{maybe skip the next sentence} 
% In fact, going beyond the class of rapidly varying distributions, (e.g., for heavy Pareto-type tails) the support recovery problem is fundamentally different and there may no longer be a phase-transition phenomenon.

%\section{Exact support recovery under AGG errors}
%\label{sec:boundary}
\input{4.exact_support_recovery/4.phase_transition_results.tex}

%\section{Bayes minimax optimality and (sub)optimality of thresholding procedures}

%\section{Strong classification boundaries in other light-tailed error models}
%\label{suppsec:other-boundaries}
%\input{4.exact_support_recovery/4.other-boundaries.tex}

%\section{Thresholding procedures under heavy-tailed errors}
%\label{suppsec:heavy-tailed}
%\input{4.exact_support_recovery/4.heavy-tailed.tex}